---
title: "rum"

format:
  html:
    code-fold: true        # Enables dropdown for code
    code-tools: true       # (Optional) Adds buttons like "Show Code"
    code-summary: "Show code"  # (Optional) Custom label for dropdown
    toc: true
    toc-location: left
    page-layout: full
editor: visual
---

# Rum

This script will run a RUM model on our data assuming a stay at home option and compare it to our result of the MDCEV.



### Data

Need to run the same data for the MDCEV which gives us final_data_2018. So going to clip the code from the file R script high cost.

The first code will produce final_data_yyyy by year and then should be able to run all the rest of the data

```{r}

library(dplyr)
library(tidyr)
library(haven)
library(readr)
library(mlogit)
library(lmtest)
library(purrr)
library(stringr)
library(readr)
library(ggplot2)
library(stargazer)
library(readxl)
library(patchwork)
library(broom)
library(tidyverse)
library(sf)
library(tigris)
library(ggrepel)
library(ggspatial)
library(ggpattern)

```

reservation_data \<- read_dta("reservation_data_merged_id_complete.dta")

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"



# Load datasets
reservation_data <- read_dta("data/reservation_data_merged_id_complete.dta")
intra_island_distance <- read.csv("data/intra_island_distance2.csv")

# Add year
reservation_data$year <- as.numeric(format(as.Date(reservation_data$CheckinDate, "%Y-%m-%d"), "%Y"))

reservation_data <- reservation_data %>%
  mutate(zip = ifelse(zip == 96812, 96813, zip))

# Group and compute yearly averages
avg_days_per_year <- reservation_data %>%
  group_by(year, id) %>%
  summarize(avg_days_visited = mean(days_visited, na.rm = TRUE)) %>%
  ungroup()

avg_days_2018 <- avg_days_per_year %>% filter(year == 2018)
avg_days_2019 <- avg_days_per_year %>% filter(year == 2019)
avg_days_2020 <- avg_days_per_year %>% filter(year == 2020)
avg_days_2021 <- avg_days_per_year %>% filter(year == 2021)
avg_days_2022 <- avg_days_per_year %>% filter(year == 2022)
avg_days_2023 <- avg_days_per_year %>% filter(year == 2023)


# Function to process data for a single year
process_year <- function(year_input) {
  data_year <- reservation_data %>% filter(year == year_input)
  
  unique_visits <- data_year %>%
    filter(!is.na(CheckinDate)) %>%
    distinct(id, parkname, CheckinDate) %>%
    group_by(id, parkname) %>%
    summarize(VisitCount = n(), .groups = "drop")
  
  all_combinations <- expand.grid(
    id = unique(data_year$id),
    parkname = unique(data_year$parkname),
    stringsAsFactors = FALSE
  )
  
  visit_data <- all_combinations %>%
    left_join(unique_visits, by = c("id", "parkname")) %>%
    mutate(VisitCount = ifelse(is.na(VisitCount), 0, VisitCount))
  
  person_info <- data_year %>%
    select(id, Permittee, zip, Zip, NoOfGuests, days_visited, year) %>%
    distinct()
  
  long_resdata_count <- left_join(visit_data, person_info, by = "id")
  
  long_resdata_count <- long_resdata_count %>%
    mutate(island_park = case_when(
      parkname == "Ainapo Cabin" ~ "Hawaii",
      parkname == "Hipalau" ~ "Kauai",
      parkname == "Kaluahaulu" ~ "Kauai",
      parkname == "Kamananui Trail" ~ "Oahu",
      parkname == "Kaunala Trail" ~ "Oahu",
      parkname == "Kawaikoi" ~ "Kauai",
      parkname == "Keanakolu Ranger & Bunkhouse Cabins" ~ "Hawaii",
      parkname == "Kuaokala Trail" ~ "Oahu",
      parkname == "Kulanaahane Trail" ~ "Oahu",
      parkname == "Kuliouou Trail" ~ "Oahu",
      parkname == "Lonomea" ~ "Kauai",
      parkname == "Maakua Ridge Trail" ~ "Oahu",
      parkname == "Manana Trail" ~ "Oahu",
      parkname == "Peacock Flats Campsite" ~ "Oahu",
      parkname == "Sugi Grove" ~ "Kauai",
      parkname == "Waialae Cabin Campsite" ~ "Kauai",
      parkname == "Waikoali" ~ "Kauai",
      parkname == "Waikolu" ~ "Molokai",
      parkname == "Waimano Trail" ~ "Oahu",
      parkname == "Waimanu Campsite" ~ "Hawaii",
      parkname == "Wiliwili" ~ "Kauai",
      parkname == "Wiliwilinui Ridge Trail" ~ "Oahu",
      TRUE ~ NA_character_
    ))
  
  long_resdata_count <- long_resdata_count[order(long_resdata_count$id, long_resdata_count$parkname), ]
  long_resdata_count$parkname_numeric <- as.integer(factor(long_resdata_count$parkname, levels = unique(long_resdata_count$parkname)))
  
  long_resdata_count <- long_resdata_count %>%
    distinct(id, parkname, .keep_all = TRUE)
  
  final_data <- long_resdata_count %>%
    left_join(
      intra_island_distance %>%
        select(zip, parkname, income, residents_distance_meters, residents_duration, island_live),
      by = c("zip", "parkname")
    )
  
  final_data <- final_data %>%
    group_by(id) %>%
    mutate(visited_another_island = ifelse(any(island_park != island_live & VisitCount > 0), 1, 0)) %>%
    ungroup()
  
  final_data$travel_distance_km <- (final_data$residents_distance_meters / 1000) * 2
  
  
  return(final_data)
}

# Run the function for each year
final_data_2018 <- process_year(2018)
final_data_2019 <- process_year(2019)
final_data_2020 <- process_year(2020)
final_data_2021 <- process_year(2021)
final_data_2022 <- process_year(2022)
final_data_2023 <- process_year(2023)

# Add correct avg_days_visited for each year by joining on id

final_data_2018 <- final_data_2018 %>%
  left_join(avg_days_2018 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2019 <- final_data_2019 %>%
  left_join(avg_days_2019 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2020 <- final_data_2020 %>%
  left_join(avg_days_2020 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2021 <- final_data_2021 %>%
  left_join(avg_days_2021 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2022 <- final_data_2022 %>%
  left_join(avg_days_2022 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2023 <- final_data_2023 %>%
  left_join(avg_days_2023 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

## Travel cost per year calculations ##


##Adjusted for inflation
gas_prices <- data.frame(
  year = c(2018, 2019, 2020, 2021, 2022, 2023),
  gas_price = c(4.68, 4.49, 4.09, 4.75, 5.81, 5.09)
)

mean_gas_price <- (4.68+4.49+4.09+4.75+5.81+5.09)/6


# Inputs based on AAA Brochure 2018
fuel_efficiency_mpg <- 25
gas_price_hawaii_2018 <- 4.68  # in USD/gallon
maintenance_per_mile <- 0.0858*1.2816  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6202*1.2816  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2018 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://exchange.aaa.com/wp-content/uploads/2018/09/18-0090_2018-Your-Driving-Costs-Brochure_FNL-Lo-5-2.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2018$price <- (final_data_2018$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2018$income * 0.75) / 2080 * final_data_2018$travel_distance_km / 50)


# Inputs based on AAA Brochure 2019
fuel_efficiency_mpg <- 25
gas_price_hawaii_2019 <- 4.49  # in USD/gallon
maintenance_per_mile <- 0.0960*1.262  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6845*1.262  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2019 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://exchange.aaa.com/wp-content/uploads/2019/09/AAA-Your-Driving-Costs-2019.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2019$price <- (final_data_2019$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2019$income * 0.75) / 2080 * final_data_2019$travel_distance_km / 50)

# Inputs based on AAA Brochure 2020
fuel_efficiency_mpg <- 25
gas_price_hawaii_2020 <- 4.09  # in USD/gallon
maintenance_per_mile <- 0.1011*1.2314  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6831*1.2314  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2020 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://newsroom.aaa.com/wp-content/uploads/2020/12/2020-Your-Driving-Costs-Brochure-Interactive-FINAL-12-9-20.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2020$price <- (final_data_2020$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2020$income * 0.75) / 2080 * final_data_2020$travel_distance_km / 50)

# Inputs based on AAA Brochure 2021
fuel_efficiency_mpg <- 25
gas_price_hawaii_2021 <- 4.75  # in USD/gallon
maintenance_per_mile <- 0.1043*1.12144  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6633*1.12144  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2021 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://newsroom.aaa.com/wp-content/uploads/2021/08/2021-YDC-Brochure-Live.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2021$price <- (final_data_2021$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2021$income * 0.75) / 2080 * final_data_2021$travel_distance_km / 50)

# Inputs based on AAA Brochure 2022
fuel_efficiency_mpg <- 25
gas_price_hawaii_2022 <- 5.81  # in USD/gallon
maintenance_per_mile <- 0.1064*1.113  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6599*1.113  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2022 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2022$price <- (final_data_2022$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2022$income * 0.75) / 2080 * final_data_2022$travel_distance_km / 50)

#https://newsroom.aaa.com/wp-content/uploads/2022/08/2022-YDC-Costs-Break-Out-by-Category.pdf

# Inputs based on AAA Brochure 2023
fuel_efficiency_mpg <- 25
gas_price_hawaii_2023 <- 5.09  # in USD/gallon
maintenance_per_mile <- 0.1085*1.0618  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 7542*1.0618  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2023 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2023$price <- (final_data_2023$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2023$income * 0.75) / 2080 * final_data_2023$travel_distance_km / 50)

#https://newsroom.aaa.com/wp-content/uploads/2023/08/YDC-Brochure_2023-FINAL-8.30.23-.pdf


final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))


final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

#########Value of travel time in air#########

### BIG ISLAND ###
##Insert value of spare time of travelling with flight to BI if live in Oahu (90 minutes before flight + 50 minutes flight time)
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu",
                                (((income * 0.75) / 2080) * 2.33) * 2, 0))

##Insert value of spare time of travelling with flight to BI if live in Kauai (90 minutes before flight + 55 minutes flight time)
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai",
                                (((income * 0.75) / 2080) * 2.42) * 2, 0))

##Insert value of spare time of travelling with flight to BI if live in Molokai (90 minutes before flight + 1.5 hour flight time one way)
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii",
                                (((income * 0.75) / 2080) * 2.33) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii",
                                (((income * 0.75) / 2080) * 2.42) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2.33) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 2) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 4.34) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 4.34) * 2, 0))


#########Value of travel time in air#########

### BIG ISLAND ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

#########Value of travel time in air#########

### BIG ISLAND ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

##

### BIG ISLAND ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

##

### BIG ISLAND ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

##

### BIG ISLAND ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))



## Flight costs 2018
#########Flight cost - 2018 #########

## BIG ISLAND ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 354.0447586, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 368.0753954, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 308.7662656, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 273.2598837, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 354.0447586 + 137.8000852, 0))  # via Oahu

## OAHU ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 354.0447586, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 269.4129138, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 112.5000815, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 291.0965957, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 137.8000852, 0))

## KAUAI ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 368.0753954, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 269.4129138, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 317.7951903, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 359.4733617, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 336.1283623, 0))

## MOLOKAI ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 308.7662656, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 112.5000815, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 317.7951903, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 118.8094464, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 137.8000852 + 112.5000815, 0))  # via Oahu

## Flight costs 2019
## BIG ISLAND ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 348.630216, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 362.446277, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 304.044185, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 269.080816, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 348.630216 + 135.692656, 0))  # via Oahu

## OAHU ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 348.630216, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 265.292679, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 110.779575, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 286.644744, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 135.692656, 0))

## KAUAI ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 362.446277, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 265.292679, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 312.935027, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 353.975798, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 330.987822, 0))

## MOLOKAI ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 304.044185, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 110.779575, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 312.935027, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 118.809446, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 135.692656 + 110.779575, 0))  # via Oahu


## Flight costs 2020
## BIG ISLAND ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 340.176901, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 353.657960, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.671957, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 262.556352, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 340.176901 + 132.402485, 0))  # via Oahu

## OAHU ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 340.176901, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 258.860067, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 108.093477, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 279.694404, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.402485, 0))

## KAUAI ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 353.657960, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 258.860067, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 305.347220, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 345.392866, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 322.962286, 0))

## MOLOKAI ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.671957, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 108.093477, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 305.347220, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 108.093477 + 279.6944039, 0), # via Oahu
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 132.402485 + 108.093477, 0))  # via Oahu

##Flight costs 2021
## Flight costs 2021
## BIG ISLAND ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 309.800214, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 322.077459, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 239.110927, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 309.800214 + 132.4024851, 0))  # via Oahu #Use value for 2020 for Oahu to Lanai

## OAHU ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 309.800214, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 235.744708, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 126.263151, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 254.718607, 0), 
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.4024851, 0)) #Use value for 2020

## KAUAI ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 322.077459, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 235.744708, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 389.441668, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 314.550411, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 235.744708 + 132.4024851, 0))  # via Oahu #Use value for 2020 for Oahu to Lanai

## MOLOKAI ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 120.699975, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 389.441668, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 120.699975 + 254.718607, 0), # via Oahu
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 126.263151 + 120.699975, 0))  # via Oahu

##Flight cost 2022
## Flight costs 2022

## BIG ISLAND ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 307.468646, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 319.6534918, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 237.311369, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 307.468646 + 132.4024851, 0))  # via Oahu

## OAHU ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 307.468646, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 233.970485, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 169.716317, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 252.801585, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.4024851, 0))  #Use value for 2020

## KAUAI ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 319.6534918, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 233.970485, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 589.133955, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 312.1830927, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 233.970485 + 132.4024851, 0))  # via Oahu #Use value for 2020

## MOLOKAI ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 169.716317, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 275.9878642, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 169.716317 + 252.8015848, 0),  # via Oahu
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 169.716317 + 132.4024851, 0))  # via Oahu

##Flight cost 2023
## HAWAII ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 293.324536, 0)) %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 304.948857, 0)) %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.6719565, 0)) %>% #Use value for 2020
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 226.39462, 0)) %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 132.4024851 + 293.324536, 0))

## OAHU ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 293.324536, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 223.207422, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 169.464607, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 241.172258, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.4024851, 0)) #Use value for 2020

## KAUAI ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 304.948857, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 223.207422, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 500.736032, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 297.822109, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 223.207422 + 132.4024851, 0))

## MOLOKAI ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.6719565, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 169.464607, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 263.291927, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 241.172258 + 169.464607, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 132.4024851 + 169.464607, 0))



#########Car rental cost#########

### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


##Mean travel cost
summary(final_data_2018$price)

final_data_2018 <- final_data_2018 %>%
  arrange(Permittee, parkname)

final_data_2018$parkname_numeric <- as.integer(factor(final_data_2018$parkname, levels = unique(final_data_2018$parkname)))

##Mean travel cost
summary(final_data_2019$price)

final_data_2019 <- final_data_2019 %>%
  arrange(Permittee, parkname)

final_data_2019$parkname_numeric <- as.integer(factor(final_data_2019$parkname, levels = unique(final_data_2019$parkname)))

##Mean travel cost
summary(final_data_2020$price)

final_data_2020 <- final_data_2020 %>%
  arrange(Permittee, parkname)

final_data_2020$parkname_numeric <- as.integer(factor(final_data_2020$parkname, levels = unique(final_data_2020$parkname)))

##Mean travel cost
summary(final_data_2021$price)

final_data_2021 <- final_data_2021 %>%
  arrange(Permittee, parkname)

final_data_2021$parkname_numeric <- as.integer(factor(final_data_2021$parkname, levels = unique(final_data_2021$parkname)))

##Mean travel cost
summary(final_data_2022$price)

final_data_2022 <- final_data_2022 %>%
  arrange(Permittee, parkname)

final_data_2022$parkname_numeric <- as.integer(factor(final_data_2022$parkname, levels = unique(final_data_2022$parkname)))

##Mean travel cost
summary(final_data_2023$price)

final_data_2023 <- final_data_2023 %>%
  arrange(Permittee, parkname)

final_data_2023$parkname_numeric <- as.integer(factor(final_data_2023$parkname, levels = unique(final_data_2023$parkname)))

```

#### Rum

```{r setup}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"


# Create unique choice occasions for every visit
data_rum <- final_data_2018 %>%
  filter(VisitCount > 0) %>%
  uncount(VisitCount) %>%
  group_by(id) %>%
  mutate(choice_id = paste0(id, "_", row_number())) %>%
  ungroup() %>%
  select(choice_id, id, chosen_park = parkname) %>%
  expand_grid(parkname = unique(final_data_2018$parkname)) %>%
  mutate(choice = (chosen_park == parkname)) %>%
  left_join(final_data_2018 %>% select(id, parkname, price/NoOfGuests, travel_distance_km,income,zip), 
            by = c("id", "parkname"))

# Create Stay at Home

# Create the Stay-at-Home option for every choice occasion
outside_home <- data_rum %>%
  distinct(choice_id, id,income,zip) %>%
  mutate(
    parkname = "Stay_Home",
    choice = FALSE,      
    price = 0,
    travel_distance_km = 0
  )

# Combine with your existing data
data_rum<- bind_rows(data_rum, outside_home)

# Re-format for mlogit
df_rum <- dfidx(data_rum, 
                        idx = list("choice_id", "parkname"), 
                        shape = "long")

# Simple Multinomial Logit
rum_mle <- mlogit(choice ~ price  | 1, data = df_rum)
summary(rum_mle)



# Calculate the 'Logsum' (Total Utility) for the status quo
logsum_base <- logsum(rum_mle)

# 1. Convert df_rum to a standard data frame for our math
df_math <- as.data.frame(df_rum)

# 2. Extract the IDs and Park names from the 'idx' column
# idx(df_rum, 1) gets the 'choice_id'
# idx(df_rum, 2) gets the 'parkname'
df_math$trip_id   <- as.character(idx(df_rum, 1))
df_math$park_name <- as.character(idx(df_rum, 2))

# 3. Calculate the Systematic Utility (V)
# This uses the model coefficients and the data matrix
df_math$V <- as.numeric(model.matrix(rum_mle) %*% coef(rum_mle))

# 4. Double check the columns are there
head(df_math[, c("trip_id", "park_name", "V")])



```

#### McFaddenR2

The mlogit gives NA so running manually to calculate McFadden to double check the R2

and it above .2 so that great

```{r}
#| cache: true
# 1. Estimate your model (The "Full" model)
rum_full <- mlogit(choice ~ price | 1, data = df_rum)

# 2. Estimate the "Null" model (Intersects only, no price)
# Use '0' for the first part of the formula to remove choice-specific variables
rum_null <- mlogit(choice ~ 0 | 1, data = df_rum)

# 3. Use the lrtest function to get your Chi-square and p-value

lr_test_result <- lrtest(rum_full, rum_null)
print(lr_test_result)

# 4. Calculate McFadden's R-squared manually
ll_full <- as.numeric(logLik(rum_full))
ll_null <- as.numeric(logLik(rum_null))
mcfadden_r2 <- 1 - (ll_full / ll_null)

print(paste("Manual McFadden R2:", round(mcfadden_r2, 4)))



```

## Function

This function calculate the rum welfare manual and is used for each year to get a comparison to the MDCEV.

# Per Person Stay-At-Home

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"


years <- 2018:2023
mcfadden_list <- list()
model_list_obs <- list()

# --- 1. Robust Welfare Function ---
# Passing 'mle' and 'data' as arguments to avoid "Global Object" confusion
calc_rum_welfare <- function(target_park, current_mle, current_math_df) {
    beta_price <- coef(current_mle)["price"]
    
    welfare <- current_math_df %>%
        group_by(trip_id) %>%
        summarise(
            ls_base = log(sum(exp(V))),
            # Calculate utility if the target park/Stay_Home was REMOVED
            ls_policy = log(sum(exp(V[park_name != target_park]))),
            .groups = "drop"
        ) %>%
        mutate(cv_trip = (ls_policy - ls_base) / -beta_price) %>%
        mutate(id = as.character(str_remove(trip_id, "_\\d+$"))) %>%
        group_by(id) %>%
        summarise(annual_rum_loss = sum(cv_trip, na.rm = TRUE), .groups = "drop")
    
    return(welfare)
}

# --- 2. Main Loop ---
final_comparison <- map_df(years, function(yr) {
    
    message(paste(">>> Processing Year:", yr))
    
    # Load data (ensure it exists)
    current_data <- get(paste0("final_data_", yr))
    
    # CLEANING: Ensure unique user profiles to avoid "Many-to-Many" warnings
    unique_users <- current_data %>% 
        group_by(id) %>%
        summarise(zip = first(zip), income = first(income), .groups = "drop") %>%
        mutate(id = as.character(id))
    
    # SITE MAPPING: Force "Stay_Home" to Site 22
    physical_parks <- unique(current_data$parkname)
    site_lookup <- data.frame(
        site = seq_along(physical_parks),
        parkname = physical_parks
    ) %>% 
    bind_rows(data.frame(site = 22, parkname = "Stay_Home"))
    
    # List of all alternatives to test (including Stay_Home)
    parks_to_test <- site_lookup$parkname
    
    # 1. Create Choice Data
    data_rum_raw <- current_data %>%
        filter(VisitCount > 0) %>%
        uncount(VisitCount) %>%
        group_by(id) %>%
        mutate(choice_id = paste0(id, "_", row_number())) %>%
        ungroup() %>%
        select(choice_id, id, chosen_park = parkname) %>%
        expand_grid(parkname = physical_parks) %>%
        mutate(choice = (chosen_park == parkname)) %>%
        left_join(current_data %>% select(id, parkname, price), by = c("id", "parkname"))
    
    # 2. Add Outside Option
    outside_home <- data_rum_raw %>%
        distinct(choice_id, id) %>%
        mutate(parkname = "Stay_Home", choice = FALSE, price = 0)
    
    data_rum_combined <- bind_rows(data_rum_raw, outside_home)
    
    # 3. Model Estimation with Error Handling (Fixes the 2020 crash)
    df_idx_local <- dfidx(data_rum_combined, idx = list("choice_id", "parkname"), shape = "long")
    
    rum_mle <- tryCatch({
        mlogit(choice ~ price | 1, data = df_idx_local, method = "nr")
    }, error = function(e) {
        # Try BHHH method if NR fails (common in 2020 data)
        message("NR failed, trying BHHH solver...")
        mlogit(choice ~ price | 1, data = df_idx_local, method = "bhhh")
    })
    
    model_list_obs[[paste0("Year_", yr)]] <<- rum_mle
    
    # 4. Prepare Math Data for this specific year
    df_math_local <- as.data.frame(df_idx_local)
    df_math_local$trip_id   <- as.character(idx(df_idx_local, 1))
    df_math_local$park_name <- as.character(idx(df_idx_local, 2))
    df_math_local$V <- as.numeric(model.matrix(rum_mle) %*% coef(rum_mle))
    
    # 5. RUM Welfare Calculation (Now includes Stay_Home)
    rum_results <- map_df(parks_to_test, function(p) {
        res <- calc_rum_welfare(p, rum_mle, df_math_local)
        res$parkname <- p
        return(res)
    })
    
    # 6. Bring in MDCEV Data (Using Site 22 for Stay_Home)
    mdcev_indiv <- read_csv(paste0("data/welfare_", yr, "_ind.csv"), show_col_types = FALSE) %>%
        left_join(site_lookup, by = "site") %>%
        mutate(id = as.character(id)) %>%
        select(id, parkname, mdcev_annual_loss = mean_value) %>%
        filter(!is.na(parkname)) # Remove sites not in our lookup
    
    # 7. Final Join
    year_comparison <- rum_results %>%
        left_join(mdcev_indiv, by = c("id", "parkname")) %>%
        left_join(unique_users, by = "id") %>% 
        mutate(
            year = yr,
            difference = annual_rum_loss - mdcev_annual_loss
        )
    
    return(year_comparison)
})

saveRDS(model_list_obs, "data/model_list_obs.rds")
```

# Per person 52-weeks

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"


years <- 2018:2023

#  Same Mcfadden and coefficient holders
model_list_52  <- list()
mcfadden_list_52 <- list()

#Welfare Function ---
calc_rum_welfare_per_person <- function(target_park, current_mle, current_math_df) {
    beta_price <- coef(current_mle)["price"]
    
    welfare_by_id <- current_math_df %>%
        group_by(trip_id) %>%
        summarise(
            ls_base = log(sum(exp(V))),
            ls_policy = log(sum(exp(V[park_name != target_park]))),
            .groups = "drop"
        ) %>%
        mutate(cv_week = (ls_policy - ls_base) / -beta_price) %>%
        mutate(id = as.character(str_remove(trip_id, "_w\\d+$"))) %>%
        group_by(id) %>%
        summarise(annual_rum_loss = sum(cv_week, na.rm = TRUE), .groups = "drop")
    
    return(welfare_by_id)
}

# --- 2. Main Loop ---
final_comparison_52_person <- map_df(years, function(yr) {
    
    message(paste(">>> Processing Year (52-week):", yr))
    current_data <- get(paste0("final_data_", yr))
    
    # Setup a 52 week decision
    unique_users <- current_data %>% 
        group_by(id) %>%
        summarise(zip = first(zip), income = first(income), .groups = "drop") %>%
        mutate(id = as.character(id))
      
    all_parks <- unique(current_data$parkname)
    
    # --- SITE LOOKUP: Force Stay_Home to Site 22 ---
    site_lookup <- data.frame(
        site = seq_along(all_parks),
        parkname = all_parks
    ) %>% 
    bind_rows(data.frame(site = 22, parkname = "Stay_Home"))
    
    # Ensure "Stay_Home" is in the welfare test list
    parks_to_test <- site_lookup$parkname
    
    # Expand to 52 Weeks
    df_52_weeks <- unique_users %>%
        expand_grid(week = 1:52) %>%
        mutate(choice_id = paste0(id, "_w", week))
    
    actual_trips <- current_data %>% 
        filter(VisitCount > 0) %>% 
        uncount(VisitCount) %>%
        group_by(id) %>%
        mutate(week = row_number()) %>%
        filter(week <= 52) %>% 
        ungroup() %>%
        select(id, week, chosen_park = parkname) %>%
        mutate(id = as.character(id))
    
    # 2. Build Choice Set
    df_rum_final <- df_52_weeks %>%
        left_join(actual_trips, by = c("id", "week")) %>%
        mutate(chosen_park = replace_na(chosen_park, "Stay_Home")) %>%
        expand_grid(parkname = c(all_parks, "Stay_Home")) %>%
        mutate(choice = (chosen_park == parkname)) %>%
        left_join(current_data %>% mutate(id = as.character(id)) %>% select(id, parkname, price), 
                  by = c("id", "parkname")) %>%
        mutate(price = replace_na(price, 0))
    
    # Model Estimation with tryCatch (Fixes 2020 error)
    df_idx_local <- dfidx(df_rum_final, idx = list("choice_id", "parkname"), shape = "long")
    
    rum_mle_52 <- tryCatch({
        mlogit(choice ~ price | 1 , data = df_idx_local, method = "nr")
    }, error = function(e) {
        message("NR Solver failed, trying BHHH...")
        mlogit(choice ~ price | 1 , data = df_idx_local, method = "bhhh")
    })

    model_list_52[[paste("52wk", yr)]] <<- rum_mle_52
    
    # McFadden R2 calculation
    rum_null <- mlogit(choice ~ 1 | 1, data = df_idx_local)
    ll_mod  <- as.numeric(logLik(rum_mle_52))
    ll_null <- as.numeric(logLik(rum_null))
    mcf_r2  <- 1 - (ll_mod / ll_null)
    mcfadden_list_52[[paste("52weeks", yr)]] <<- format(round(mcf_r2, 3), nsmall = 3)

    df_math_local <- as.data.frame(df_idx_local)
    df_math_local$trip_id   <- as.character(idx(df_idx_local, 1))
    df_math_local$park_name <- as.character(idx(df_idx_local, 2))
    df_math_local$V <- as.numeric(model.matrix(rum_mle_52) %*% coef(rum_mle_52))
    
    # RUM Welfare Calculations (Including Stay_Home)
    rum_person_results <- map_df(parks_to_test, function(p) {
        res <- calc_rum_welfare_per_person(p, rum_mle_52, df_math_local)
        res$parkname <- p
        return(res)
    })
    
    # Bring in MDCEV Individual Welfare
    mdcev_indiv <- read_csv(paste0("data/welfare_", yr, "_ind.csv"), show_col_types = FALSE) %>%
        left_join(site_lookup, by = "site") %>%
        mutate(id = as.character(id)) %>%
        select(id, parkname, mdcev_annual_loss = mean_value) %>%
        filter(!is.na(parkname)) # Ensures we only keep matched sites
    
    # Final Join
    year_comparison <- rum_person_results %>%
        left_join(mdcev_indiv, by = c("id", "parkname")) %>%
        left_join(unique_users, by = "id") %>% 
        mutate(
            year = yr,
            difference = annual_rum_loss - mdcev_annual_loss
        )
    
    return(year_comparison)
})

saveRDS(model_list_52, "data/model_list_52.rds")


```

# Full Closure

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"
# 1. Extract the Price Coefficient
beta_price <- coef(rum_mle)["price"]

# 2. Calculate Total System Welfare Loss
# We compare the current state (All Parks) to the "Closed" state (Stay_Home Only)
total_system_welfare <- df_math %>%
  group_by(trip_id) %>%
  summarise(
    # Inclusive value (Log-Sum) of the current world:
    ls_base = log(sum(exp(V))),
    
    # Inclusive value if ONLY 'Stay_Home' exists (Everything else removed):
    ls_policy = log(sum(exp(V[park_name == "Stay_Home"]))),
    
    .groups = "drop"
  ) %>%
  # Welfare formula: (LS_policy - LS_base) / -beta_price
  mutate(cv_week = (ls_policy - ls_base) / -beta_price) %>%
  
  # Group by User ID to get the Annual Total
  mutate(id = as.character(str_remove(trip_id, "_w\\d+$"))) %>%
  group_by(id) %>%
  summarise(annual_system_loss = sum(cv_week, na.rm = TRUE), .groups = "drop")

# 3. View the Results
summary_stats <- total_system_welfare %>%
  summarise(
    Mean_Annual_Value = mean(annual_system_loss),
    Median_Annual_Value = median(annual_system_loss),
    Total_Sample_Value = sum(annual_system_loss)
  )

print(summary_stats)

```

# Coeffienct Table

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"

# Create a specific list for the 2018/2023 comparison
# We alternate them: [Obs 2018, 52wk 2018, Obs 2023, 52wk 2023]
comparison_list <- list(
    "Observed" = model_list_obs[["Year_2018"]], 
    "52-Week"  = model_list_52[["52wk 2018"]],
    "Observed" = model_list_obs[["Year_2023"]], 
    "52-Week"  = model_list_52[["52wk 2023"]]
)
```

```{r}
#| cache: true
#| eval: false
# 2. Generate the Professionally Grouped Table

# Ensure your lists exist and are not empty before running this
stargazer(
  comparison_list,
  type = "html",
  out = "tables/rum_comparison_results.html",
  title = "Table: Comparison of RUM Specifications (2018 vs. 2023)",
  
  # --- THE KEY ARGUMENT FOR SIDE-BY-SIDE SEs ---
  single.row = TRUE, 
  
  column.labels = c("2018", "2023"),
  column.separate = c(2, 2),
  model.names = TRUE,
  dep.var.labels.include = FALSE,
  model.numbers = FALSE,
  
  # Pulling Price (22) and Bunker (6) to the top
  order = c(22, 6),
  covariate.labels = c("Travel Cost (Price)", "Keanakolu Bunkhouse"),
  
  omit.stat = c("f", "ser", "lr", "rsq", "adj.rsq"),
  star.cutoffs = c(0.05, 0.01, 0.001),
  no.space = TRUE,
  notes = "Standard errors in parentheses next to coefficients. 52-Week model includes 'Stay-at-Home'."
)

```

# All Coefficent

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"

# Create a specific list for the 2018/2023 comparison

comparison_list <- list(
    "Observed" = model_list_obs[["Year_2018"]], 
    "52-Week"  = model_list_52[["52wk 2018"]],
    "Observed" = model_list_obs[["Year_2019"]], 
    "52-Week"  = model_list_52[["52wk 2019"]],
    "Observed" = model_list_obs[["Year_2020"]], 
    "52-Week"  = model_list_52[["52wk 2020"]],
    "Observed" = model_list_obs[["Year_2021"]], 
    "52-Week"  = model_list_52[["52wk 2021"]],
    "Observed" = model_list_obs[["Year_2022"]], 
    "52-Week"  = model_list_52[["52wk 2022"]],
    "Observed" = model_list_obs[["Year_2023"]], 
    "52-Week"  = model_list_52[["52wk 2023"]]
)

```

```{r}
#| cache: true
#| eval: false

stargazer(
comparison_list,
type = "html",
out = "tables/rum_comparison_results_appendix.html",
title = "Table: Comparison of RUM Specifications (2018 vs. 2023)",
column.labels = c("2018", "2019","2020","2021","2022","2023"),
column.separate = c(2, 2),
model.names = TRUE,
dep.var.labels.include = FALSE,
model.numbers = FALSE,
# Use 'order' to pull Price (22) and Bunker (6) to the top of the table
# This keeps the table concise and professional
order = c(23, 6),
# Rename only the ones we kept at the top
covariate.labels = c("Travel Cost (Price)", "Keanakolu Bunkhouse"),
omit.stat = c("f", "ser", "lr", "rsq", "adj.rsq"),
star.cutoffs = c(0.05, 0.01, 0.001),
no.space = TRUE,
notes = "Standard errors in parentheses. 52-Week model includes 'Stay-at-Home' as a weekly choice."
)

```

#### Merge

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"


final_comparison_master <- final_comparison %>%
  # Select and rename columns from the first dataset
  select(id, parkname, year, zip, income, 
         loss_rum_observed = annual_rum_loss, 
         loss_mdcev = mdcev_annual_loss) %>%
  

  full_join(
    final_comparison_52_person %>% 
      select(id, parkname, year, loss_rum_52wk = annual_rum_loss), 
    by = c("id", "parkname", "year")
  ) 

head(final_comparison_master)

#write_csv(final_comparison_master, "data/final_comparison_master.csv")

```
