---
title: "rum"

format:
  html:
    code-fold: true        # Enables dropdown for code
    code-tools: true       # (Optional) Adds buttons like "Show Code"
    code-summary: "Show code"  # (Optional) Custom label for dropdown
    toc: true
    toc-location: left
    page-layout: full
editor: visual
---

# Rum

This script will run a RUM model on our data assuming a stay at home option and compare it to our result of the MDCEV

### Data

Need to run the same data for the MDCEV which gives us final_data_2018. So going to clip the code from the file R script high cost.

The first code will produce final_data_yyyy by year and then should be able to run all the rest of the data

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"

library(dplyr)
library(tidyr)
 library(haven)
library(readr)

# Load datasets
reservation_data <- read_dta("reservation_data_merged_id_complete.dta")
intra_island_distance <- read.csv("intra_island_distance2.csv")

# Add year
reservation_data$year <- as.numeric(format(as.Date(reservation_data$CheckinDate, "%Y-%m-%d"), "%Y"))

reservation_data <- reservation_data %>%
  mutate(zip = ifelse(zip == 96812, 96813, zip))

# Group and compute yearly averages
avg_days_per_year <- reservation_data %>%
  group_by(year, id) %>%
  summarize(avg_days_visited = mean(days_visited, na.rm = TRUE)) %>%
  ungroup()

avg_days_2018 <- avg_days_per_year %>% filter(year == 2018)
avg_days_2019 <- avg_days_per_year %>% filter(year == 2019)
avg_days_2020 <- avg_days_per_year %>% filter(year == 2020)
avg_days_2021 <- avg_days_per_year %>% filter(year == 2021)
avg_days_2022 <- avg_days_per_year %>% filter(year == 2022)
avg_days_2023 <- avg_days_per_year %>% filter(year == 2023)


# Function to process data for a single year
process_year <- function(year_input) {
  data_year <- reservation_data %>% filter(year == year_input)
  
  unique_visits <- data_year %>%
    filter(!is.na(CheckinDate)) %>%
    distinct(id, parkname, CheckinDate) %>%
    group_by(id, parkname) %>%
    summarize(VisitCount = n(), .groups = "drop")
  
  all_combinations <- expand.grid(
    id = unique(data_year$id),
    parkname = unique(data_year$parkname),
    stringsAsFactors = FALSE
  )
  
  visit_data <- all_combinations %>%
    left_join(unique_visits, by = c("id", "parkname")) %>%
    mutate(VisitCount = ifelse(is.na(VisitCount), 0, VisitCount))
  
  person_info <- data_year %>%
    select(id, Permittee, zip, Zip, NoOfGuests, days_visited, year) %>%
    distinct()
  
  long_resdata_count <- left_join(visit_data, person_info, by = "id")
  
  long_resdata_count <- long_resdata_count %>%
    mutate(island_park = case_when(
      parkname == "Ainapo Cabin" ~ "Hawaii",
      parkname == "Hipalau" ~ "Kauai",
      parkname == "Kaluahaulu" ~ "Kauai",
      parkname == "Kamananui Trail" ~ "Oahu",
      parkname == "Kaunala Trail" ~ "Oahu",
      parkname == "Kawaikoi" ~ "Kauai",
      parkname == "Keanakolu Ranger & Bunkhouse Cabins" ~ "Hawaii",
      parkname == "Kuaokala Trail" ~ "Oahu",
      parkname == "Kulanaahane Trail" ~ "Oahu",
      parkname == "Kuliouou Trail" ~ "Oahu",
      parkname == "Lonomea" ~ "Kauai",
      parkname == "Maakua Ridge Trail" ~ "Oahu",
      parkname == "Manana Trail" ~ "Oahu",
      parkname == "Peacock Flats Campsite" ~ "Oahu",
      parkname == "Sugi Grove" ~ "Kauai",
      parkname == "Waialae Cabin Campsite" ~ "Kauai",
      parkname == "Waikoali" ~ "Kauai",
      parkname == "Waikolu" ~ "Molokai",
      parkname == "Waimano Trail" ~ "Oahu",
      parkname == "Waimanu Campsite" ~ "Hawaii",
      parkname == "Wiliwili" ~ "Kauai",
      parkname == "Wiliwilinui Ridge Trail" ~ "Oahu",
      TRUE ~ NA_character_
    ))
  
  long_resdata_count <- long_resdata_count[order(long_resdata_count$id, long_resdata_count$parkname), ]
  long_resdata_count$parkname_numeric <- as.integer(factor(long_resdata_count$parkname, levels = unique(long_resdata_count$parkname)))
  
  long_resdata_count <- long_resdata_count %>%
    distinct(id, parkname, .keep_all = TRUE)
  
  final_data <- long_resdata_count %>%
    left_join(
      intra_island_distance %>%
        select(zip, parkname, income, residents_distance_meters, residents_duration, island_live),
      by = c("zip", "parkname")
    )
  
  final_data <- final_data %>%
    group_by(id) %>%
    mutate(visited_another_island = ifelse(any(island_park != island_live & VisitCount > 0), 1, 0)) %>%
    ungroup()
  
  final_data$travel_distance_km <- (final_data$residents_distance_meters / 1000) * 2
  
  
  return(final_data)
}

# Run the function for each year
final_data_2018 <- process_year(2018)
final_data_2019 <- process_year(2019)
final_data_2020 <- process_year(2020)
final_data_2021 <- process_year(2021)
final_data_2022 <- process_year(2022)
final_data_2023 <- process_year(2023)

# Add correct avg_days_visited for each year by joining on id

final_data_2018 <- final_data_2018 %>%
  left_join(avg_days_2018 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2019 <- final_data_2019 %>%
  left_join(avg_days_2019 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2020 <- final_data_2020 %>%
  left_join(avg_days_2020 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2021 <- final_data_2021 %>%
  left_join(avg_days_2021 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2022 <- final_data_2022 %>%
  left_join(avg_days_2022 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2023 <- final_data_2023 %>%
  left_join(avg_days_2023 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

## Travel cost per year calculations ##


##Adjusted for inflation
gas_prices <- data.frame(
  year = c(2018, 2019, 2020, 2021, 2022, 2023),
  gas_price = c(4.68, 4.49, 4.09, 4.75, 5.81, 5.09)
)

mean_gas_price <- (4.68+4.49+4.09+4.75+5.81+5.09)/6


# Inputs based on AAA Brochure 2018
fuel_efficiency_mpg <- 25
gas_price_hawaii_2018 <- 4.68  # in USD/gallon
maintenance_per_mile <- 0.0858*1.2816  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6202*1.2816  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2018 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://exchange.aaa.com/wp-content/uploads/2018/09/18-0090_2018-Your-Driving-Costs-Brochure_FNL-Lo-5-2.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2018$price <- (final_data_2018$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2018$income * 0.75) / 2080 * final_data_2018$travel_distance_km / 50)


# Inputs based on AAA Brochure 2019
fuel_efficiency_mpg <- 25
gas_price_hawaii_2019 <- 4.49  # in USD/gallon
maintenance_per_mile <- 0.0960*1.262  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6845*1.262  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2019 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://exchange.aaa.com/wp-content/uploads/2019/09/AAA-Your-Driving-Costs-2019.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2019$price <- (final_data_2019$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2019$income * 0.75) / 2080 * final_data_2019$travel_distance_km / 50)

# Inputs based on AAA Brochure 2020
fuel_efficiency_mpg <- 25
gas_price_hawaii_2020 <- 4.09  # in USD/gallon
maintenance_per_mile <- 0.1011*1.2314  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6831*1.2314  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2020 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://newsroom.aaa.com/wp-content/uploads/2020/12/2020-Your-Driving-Costs-Brochure-Interactive-FINAL-12-9-20.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2020$price <- (final_data_2020$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2020$income * 0.75) / 2080 * final_data_2020$travel_distance_km / 50)

# Inputs based on AAA Brochure 2021
fuel_efficiency_mpg <- 25
gas_price_hawaii_2021 <- 4.75  # in USD/gallon
maintenance_per_mile <- 0.1043*1.12144  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6633*1.12144  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2021 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://newsroom.aaa.com/wp-content/uploads/2021/08/2021-YDC-Brochure-Live.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2021$price <- (final_data_2021$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2021$income * 0.75) / 2080 * final_data_2021$travel_distance_km / 50)

# Inputs based on AAA Brochure 2022
fuel_efficiency_mpg <- 25
gas_price_hawaii_2022 <- 5.81  # in USD/gallon
maintenance_per_mile <- 0.1064*1.113  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6599*1.113  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2022 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2022$price <- (final_data_2022$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2022$income * 0.75) / 2080 * final_data_2022$travel_distance_km / 50)

#https://newsroom.aaa.com/wp-content/uploads/2022/08/2022-YDC-Costs-Break-Out-by-Category.pdf

# Inputs based on AAA Brochure 2023
fuel_efficiency_mpg <- 25
gas_price_hawaii_2023 <- 5.09  # in USD/gallon
maintenance_per_mile <- 0.1085*1.0618  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 7542*1.0618  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2023 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2023$price <- (final_data_2023$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2023$income * 0.75) / 2080 * final_data_2023$travel_distance_km / 50)

#https://newsroom.aaa.com/wp-content/uploads/2023/08/YDC-Brochure_2023-FINAL-8.30.23-.pdf


final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))


final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

#########Value of travel time in air#########

### BIG ISLAND ###
##Insert value of spare time of travelling with flight to BI if live in Oahu (90 minutes before flight + 50 minutes flight time)
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu",
                                (((income * 0.75) / 2080) * 2.33) * 2, 0))

##Insert value of spare time of travelling with flight to BI if live in Kauai (90 minutes before flight + 55 minutes flight time)
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai",
                                (((income * 0.75) / 2080) * 2.42) * 2, 0))

##Insert value of spare time of travelling with flight to BI if live in Molokai (90 minutes before flight + 1.5 hour flight time one way)
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii",
                                (((income * 0.75) / 2080) * 2.33) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii",
                                (((income * 0.75) / 2080) * 2.42) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2.33) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 2) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 4.34) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 4.34) * 2, 0))


#########Value of travel time in air#########

### BIG ISLAND ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

#########Value of travel time in air#########

### BIG ISLAND ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

##

### BIG ISLAND ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

##

### BIG ISLAND ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

##

### BIG ISLAND ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))



## Flight costs 2018
#########Flight cost - 2018 #########

## BIG ISLAND ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 354.0447586, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 368.0753954, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 308.7662656, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 273.2598837, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 354.0447586 + 137.8000852, 0))  # via Oahu

## OAHU ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 354.0447586, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 269.4129138, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 112.5000815, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 291.0965957, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 137.8000852, 0))

## KAUAI ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 368.0753954, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 269.4129138, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 317.7951903, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 359.4733617, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 336.1283623, 0))

## MOLOKAI ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 308.7662656, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 112.5000815, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 317.7951903, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 118.8094464, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 137.8000852 + 112.5000815, 0))  # via Oahu

## Flight costs 2019
## BIG ISLAND ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 348.630216, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 362.446277, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 304.044185, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 269.080816, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 348.630216 + 135.692656, 0))  # via Oahu

## OAHU ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 348.630216, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 265.292679, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 110.779575, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 286.644744, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 135.692656, 0))

## KAUAI ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 362.446277, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 265.292679, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 312.935027, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 353.975798, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 330.987822, 0))

## MOLOKAI ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 304.044185, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 110.779575, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 312.935027, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 118.809446, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 135.692656 + 110.779575, 0))  # via Oahu


## Flight costs 2020
## BIG ISLAND ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 340.176901, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 353.657960, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.671957, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 262.556352, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 340.176901 + 132.402485, 0))  # via Oahu

## OAHU ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 340.176901, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 258.860067, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 108.093477, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 279.694404, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.402485, 0))

## KAUAI ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 353.657960, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 258.860067, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 305.347220, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 345.392866, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 322.962286, 0))

## MOLOKAI ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.671957, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 108.093477, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 305.347220, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 108.093477 + 279.6944039, 0), # via Oahu
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 132.402485 + 108.093477, 0))  # via Oahu

##Flight costs 2021
## Flight costs 2021
## BIG ISLAND ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 309.800214, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 322.077459, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 239.110927, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 309.800214 + 132.4024851, 0))  # via Oahu #Use value for 2020 for Oahu to Lanai

## OAHU ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 309.800214, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 235.744708, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 126.263151, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 254.718607, 0), 
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.4024851, 0)) #Use value for 2020

## KAUAI ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 322.077459, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 235.744708, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 389.441668, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 314.550411, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 235.744708 + 132.4024851, 0))  # via Oahu #Use value for 2020 for Oahu to Lanai

## MOLOKAI ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 120.699975, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 389.441668, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 120.699975 + 254.718607, 0), # via Oahu
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 126.263151 + 120.699975, 0))  # via Oahu

##Flight cost 2022
## Flight costs 2022

## BIG ISLAND ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 307.468646, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 319.6534918, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 237.311369, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 307.468646 + 132.4024851, 0))  # via Oahu

## OAHU ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 307.468646, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 233.970485, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 169.716317, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 252.801585, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.4024851, 0))  #Use value for 2020

## KAUAI ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 319.6534918, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 233.970485, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 589.133955, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 312.1830927, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 233.970485 + 132.4024851, 0))  # via Oahu #Use value for 2020

## MOLOKAI ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 169.716317, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 275.9878642, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 169.716317 + 252.8015848, 0),  # via Oahu
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 169.716317 + 132.4024851, 0))  # via Oahu

##Flight cost 2023
## HAWAII ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 293.324536, 0)) %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 304.948857, 0)) %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.6719565, 0)) %>% #Use value for 2020
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 226.39462, 0)) %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 132.4024851 + 293.324536, 0))

## OAHU ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 293.324536, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 223.207422, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 169.464607, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 241.172258, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.4024851, 0)) #Use value for 2020

## KAUAI ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 304.948857, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 223.207422, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 500.736032, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 297.822109, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 223.207422 + 132.4024851, 0))

## MOLOKAI ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.6719565, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 169.464607, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 263.291927, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 241.172258 + 169.464607, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 132.4024851 + 169.464607, 0))



#########Car rental cost#########

### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


##Mean travel cost
summary(final_data_2018$price)

final_data_2018 <- final_data_2018 %>%
  arrange(Permittee, parkname)

final_data_2018$parkname_numeric <- as.integer(factor(final_data_2018$parkname, levels = unique(final_data_2018$parkname)))

##Mean travel cost
summary(final_data_2019$price)

final_data_2019 <- final_data_2019 %>%
  arrange(Permittee, parkname)

final_data_2019$parkname_numeric <- as.integer(factor(final_data_2019$parkname, levels = unique(final_data_2019$parkname)))

##Mean travel cost
summary(final_data_2020$price)

final_data_2020 <- final_data_2020 %>%
  arrange(Permittee, parkname)

final_data_2020$parkname_numeric <- as.integer(factor(final_data_2020$parkname, levels = unique(final_data_2020$parkname)))

##Mean travel cost
summary(final_data_2021$price)

final_data_2021 <- final_data_2021 %>%
  arrange(Permittee, parkname)

final_data_2021$parkname_numeric <- as.integer(factor(final_data_2021$parkname, levels = unique(final_data_2021$parkname)))

##Mean travel cost
summary(final_data_2022$price)

final_data_2022 <- final_data_2022 %>%
  arrange(Permittee, parkname)

final_data_2022$parkname_numeric <- as.integer(factor(final_data_2022$parkname, levels = unique(final_data_2022$parkname)))

##Mean travel cost
summary(final_data_2023$price)

final_data_2023 <- final_data_2023 %>%
  arrange(Permittee, parkname)

final_data_2023$parkname_numeric <- as.integer(factor(final_data_2023$parkname, levels = unique(final_data_2023$parkname)))

```

#### Rum

```{r setup}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"
library(mlogit)


# Create unique choice occasions for every visit
data_rum <- final_data_2018 %>%
  filter(VisitCount > 0) %>%
  uncount(VisitCount) %>%
  group_by(id) %>%
  mutate(choice_id = paste0(id, "_", row_number())) %>%
  ungroup() %>%
  select(choice_id, id, chosen_park = parkname) %>%
  expand_grid(parkname = unique(final_data_2018$parkname)) %>%
  mutate(choice = (chosen_park == parkname)) %>%
  left_join(final_data_2018 %>% select(id, parkname, price/NoOfGuests, travel_distance_km,income,zip), 
            by = c("id", "parkname"))

# Create Stay at Home

# Create the Stay-at-Home option for every choice occasion
outside_home <- data_rum %>%
  distinct(choice_id, id,income,zip) %>%
  mutate(
    parkname = "Stay_Home",
    choice = FALSE,      
    price = 0,
    travel_distance_km = 0
  )

# Combine with your existing data
data_rum<- bind_rows(data_rum, outside_home)

# Re-format for mlogit
df_rum <- dfidx(data_rum, 
                        idx = list("choice_id", "parkname"), 
                        shape = "long")

# Simple Multinomial Logit
rum_mle <- mlogit(choice ~ price  | 1, data = df_rum)
summary(rum_mle)



# Calculate the 'Logsum' (Total Utility) for the status quo
logsum_base <- logsum(rum_mle)

# 1. Convert df_rum to a standard data frame for our math
df_math <- as.data.frame(df_rum)

# 2. Extract the IDs and Park names from the 'idx' column
# idx(df_rum, 1) gets the 'choice_id'
# idx(df_rum, 2) gets the 'parkname'
df_math$trip_id   <- as.character(idx(df_rum, 1))
df_math$park_name <- as.character(idx(df_rum, 2))

# 3. Calculate the Systematic Utility (V)
# This uses the model coefficients and the data matrix
df_math$V <- as.numeric(model.matrix(rum_mle) %*% coef(rum_mle))

# 4. Double check the columns are there
head(df_math[, c("trip_id", "park_name", "V")])



```

#### McFaddenR2

The mlogit gives NA so running manually to calculate McFadden to double check the R2

and it above .2 so that great

```{r}
#| cache: true
# 1. Estimate your model (The "Full" model)
rum_full <- mlogit(choice ~ price | 1, data = df_rum)

# 2. Estimate the "Null" model (Intersects only, no price)
# Use '0' for the first part of the formula to remove choice-specific variables
rum_null <- mlogit(choice ~ 0 | 1, data = df_rum)

# 3. Use the lrtest function to get your Chi-square and p-value
library(lmtest)
lr_test_result <- lrtest(rum_full, rum_null)
print(lr_test_result)

# 4. Calculate McFadden's R-squared manually
ll_full <- as.numeric(logLik(rum_full))
ll_null <- as.numeric(logLik(rum_null))
mcfadden_r2 <- 1 - (ll_full / ll_null)

print(paste("Manual McFadden R2:", round(mcfadden_r2, 4)))



```

## Function

This function calculate the rum welfare manual and is used for each year to get a comparison to the MDCEV.

# Per Person Stay-At-Home

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"
library(dplyr)
library(tidyr)
library(purrr)
library(mlogit)
library(stringr)
library(readr)

years <- 2018:2023
mcfadden_list <- list()
model_list_obs <- list()

# --- 1. Robust Welfare Function ---
# Passing 'mle' and 'data' as arguments to avoid "Global Object" confusion
calc_rum_welfare <- function(target_park, current_mle, current_math_df) {
    beta_price <- coef(current_mle)["price"]
    
    welfare <- current_math_df %>%
        group_by(trip_id) %>%
        summarise(
            ls_base = log(sum(exp(V))),
            # Calculate utility if the target park/Stay_Home was REMOVED
            ls_policy = log(sum(exp(V[park_name != target_park]))),
            .groups = "drop"
        ) %>%
        mutate(cv_trip = (ls_policy - ls_base) / -beta_price) %>%
        mutate(id = as.character(str_remove(trip_id, "_\\d+$"))) %>%
        group_by(id) %>%
        summarise(annual_rum_loss = sum(cv_trip, na.rm = TRUE), .groups = "drop")
    
    return(welfare)
}

# --- 2. Main Loop ---
final_comparison <- map_df(years, function(yr) {
    
    message(paste(">>> Processing Year:", yr))
    
    # Load data (ensure it exists)
    current_data <- get(paste0("final_data_", yr))
    
    # CLEANING: Ensure unique user profiles to avoid "Many-to-Many" warnings
    unique_users <- current_data %>% 
        group_by(id) %>%
        summarise(zip = first(zip), income = first(income), .groups = "drop") %>%
        mutate(id = as.character(id))
    
    # SITE MAPPING: Force "Stay_Home" to Site 22
    physical_parks <- unique(current_data$parkname)
    site_lookup <- data.frame(
        site = seq_along(physical_parks),
        parkname = physical_parks
    ) %>% 
    bind_rows(data.frame(site = 22, parkname = "Stay_Home"))
    
    # List of all alternatives to test (including Stay_Home)
    parks_to_test <- site_lookup$parkname
    
    # 1. Create Choice Data
    data_rum_raw <- current_data %>%
        filter(VisitCount > 0) %>%
        uncount(VisitCount) %>%
        group_by(id) %>%
        mutate(choice_id = paste0(id, "_", row_number())) %>%
        ungroup() %>%
        select(choice_id, id, chosen_park = parkname) %>%
        expand_grid(parkname = physical_parks) %>%
        mutate(choice = (chosen_park == parkname)) %>%
        left_join(current_data %>% select(id, parkname, price), by = c("id", "parkname"))
    
    # 2. Add Outside Option
    outside_home <- data_rum_raw %>%
        distinct(choice_id, id) %>%
        mutate(parkname = "Stay_Home", choice = FALSE, price = 0)
    
    data_rum_combined <- bind_rows(data_rum_raw, outside_home)
    
    # 3. Model Estimation with Error Handling (Fixes the 2020 crash)
    df_idx_local <- dfidx(data_rum_combined, idx = list("choice_id", "parkname"), shape = "long")
    
    rum_mle <- tryCatch({
        mlogit(choice ~ price | 1, data = df_idx_local, method = "nr")
    }, error = function(e) {
        # Try BHHH method if NR fails (common in 2020 data)
        message("NR failed, trying BHHH solver...")
        mlogit(choice ~ price | 1, data = df_idx_local, method = "bhhh")
    })
    
    model_list_obs[[paste0("Year_", yr)]] <<- rum_mle
    
    # 4. Prepare Math Data for this specific year
    df_math_local <- as.data.frame(df_idx_local)
    df_math_local$trip_id   <- as.character(idx(df_idx_local, 1))
    df_math_local$park_name <- as.character(idx(df_idx_local, 2))
    df_math_local$V <- as.numeric(model.matrix(rum_mle) %*% coef(rum_mle))
    
    # 5. RUM Welfare Calculation (Now includes Stay_Home)
    rum_results <- map_df(parks_to_test, function(p) {
        res <- calc_rum_welfare(p, rum_mle, df_math_local)
        res$parkname <- p
        return(res)
    })
    
    # 6. Bring in MDCEV Data (Using Site 22 for Stay_Home)
    mdcev_indiv <- read_csv(paste0("welfare_", yr, "_ind.csv"), show_col_types = FALSE) %>%
        left_join(site_lookup, by = "site") %>%
        mutate(id = as.character(id)) %>%
        select(id, parkname, mdcev_annual_loss = mean_value) %>%
        filter(!is.na(parkname)) # Remove sites not in our lookup
    
    # 7. Final Join
    year_comparison <- rum_results %>%
        left_join(mdcev_indiv, by = c("id", "parkname")) %>%
        left_join(unique_users, by = "id") %>% 
        mutate(
            year = yr,
            difference = annual_rum_loss - mdcev_annual_loss
        )
    
    return(year_comparison)
})
```

# Per person 52-weeks

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"
library(dplyr)
library(tidyr)
library(purrr)
library(mlogit)
library(stringr)
library(readr)
library(ggplot2)

years <- 2018:2023

#  Same Mcfadden and coefficient holders
model_list_52  <- list()
mcfadden_list_52 <- list()

#Welfare Function ---
calc_rum_welfare_per_person <- function(target_park, current_mle, current_math_df) {
    beta_price <- coef(current_mle)["price"]
    
    welfare_by_id <- current_math_df %>%
        group_by(trip_id) %>%
        summarise(
            ls_base = log(sum(exp(V))),
            ls_policy = log(sum(exp(V[park_name != target_park]))),
            .groups = "drop"
        ) %>%
        mutate(cv_week = (ls_policy - ls_base) / -beta_price) %>%
        mutate(id = as.character(str_remove(trip_id, "_w\\d+$"))) %>%
        group_by(id) %>%
        summarise(annual_rum_loss = sum(cv_week, na.rm = TRUE), .groups = "drop")
    
    return(welfare_by_id)
}

# --- 2. Main Loop ---
final_comparison_52_person <- map_df(years, function(yr) {
    
    message(paste(">>> Processing Year (52-week):", yr))
    current_data <- get(paste0("final_data_", yr))
    
    # Setup a 52 week decision
    unique_users <- current_data %>% 
        group_by(id) %>%
        summarise(zip = first(zip), income = first(income), .groups = "drop") %>%
        mutate(id = as.character(id))
      
    all_parks <- unique(current_data$parkname)
    
    # --- SITE LOOKUP: Force Stay_Home to Site 22 ---
    site_lookup <- data.frame(
        site = seq_along(all_parks),
        parkname = all_parks
    ) %>% 
    bind_rows(data.frame(site = 22, parkname = "Stay_Home"))
    
    # Ensure "Stay_Home" is in the welfare test list
    parks_to_test <- site_lookup$parkname
    
    # Expand to 52 Weeks
    df_52_weeks <- unique_users %>%
        expand_grid(week = 1:52) %>%
        mutate(choice_id = paste0(id, "_w", week))
    
    actual_trips <- current_data %>% 
        filter(VisitCount > 0) %>% 
        uncount(VisitCount) %>%
        group_by(id) %>%
        mutate(week = row_number()) %>%
        filter(week <= 52) %>% 
        ungroup() %>%
        select(id, week, chosen_park = parkname) %>%
        mutate(id = as.character(id))
    
    # 2. Build Choice Set
    df_rum_final <- df_52_weeks %>%
        left_join(actual_trips, by = c("id", "week")) %>%
        mutate(chosen_park = replace_na(chosen_park, "Stay_Home")) %>%
        expand_grid(parkname = c(all_parks, "Stay_Home")) %>%
        mutate(choice = (chosen_park == parkname)) %>%
        left_join(current_data %>% mutate(id = as.character(id)) %>% select(id, parkname, price), 
                  by = c("id", "parkname")) %>%
        mutate(price = replace_na(price, 0))
    
    # Model Estimation with tryCatch (Fixes 2020 error)
    df_idx_local <- dfidx(df_rum_final, idx = list("choice_id", "parkname"), shape = "long")
    
    rum_mle_52 <- tryCatch({
        mlogit(choice ~ price | 1 , data = df_idx_local, method = "nr")
    }, error = function(e) {
        message("NR Solver failed, trying BHHH...")
        mlogit(choice ~ price | 1 , data = df_idx_local, method = "bhhh")
    })

    model_list_52[[paste("52wk", yr)]] <<- rum_mle_52
    
    # McFadden R2 calculation
    rum_null <- mlogit(choice ~ 1 | 1, data = df_idx_local)
    ll_mod  <- as.numeric(logLik(rum_mle_52))
    ll_null <- as.numeric(logLik(rum_null))
    mcf_r2  <- 1 - (ll_mod / ll_null)
    mcfadden_list_52[[paste("52weeks", yr)]] <<- format(round(mcf_r2, 3), nsmall = 3)

    df_math_local <- as.data.frame(df_idx_local)
    df_math_local$trip_id   <- as.character(idx(df_idx_local, 1))
    df_math_local$park_name <- as.character(idx(df_idx_local, 2))
    df_math_local$V <- as.numeric(model.matrix(rum_mle_52) %*% coef(rum_mle_52))
    
    # RUM Welfare Calculations (Including Stay_Home)
    rum_person_results <- map_df(parks_to_test, function(p) {
        res <- calc_rum_welfare_per_person(p, rum_mle_52, df_math_local)
        res$parkname <- p
        return(res)
    })
    
    # Bring in MDCEV Individual Welfare
    mdcev_indiv <- read_csv(paste0("welfare_", yr, "_ind.csv"), show_col_types = FALSE) %>%
        left_join(site_lookup, by = "site") %>%
        mutate(id = as.character(id)) %>%
        select(id, parkname, mdcev_annual_loss = mean_value) %>%
        filter(!is.na(parkname)) # Ensures we only keep matched sites
    
    # Final Join
    year_comparison <- rum_person_results %>%
        left_join(mdcev_indiv, by = c("id", "parkname")) %>%
        left_join(unique_users, by = "id") %>% 
        mutate(
            year = yr,
            difference = annual_rum_loss - mdcev_annual_loss
        )
    
    return(year_comparison)
})


```

# Full Closure

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"
# 1. Extract the Price Coefficient
beta_price <- coef(rum_mle)["price"]

# 2. Calculate Total System Welfare Loss
# We compare the current state (All Parks) to the "Closed" state (Stay_Home Only)
total_system_welfare <- df_math %>%
  group_by(trip_id) %>%
  summarise(
    # Inclusive value (Log-Sum) of the current world:
    ls_base = log(sum(exp(V))),
    
    # Inclusive value if ONLY 'Stay_Home' exists (Everything else removed):
    ls_policy = log(sum(exp(V[park_name == "Stay_Home"]))),
    
    .groups = "drop"
  ) %>%
  # Welfare formula: (LS_policy - LS_base) / -beta_price
  mutate(cv_week = (ls_policy - ls_base) / -beta_price) %>%
  
  # Group by User ID to get the Annual Total
  mutate(id = as.character(str_remove(trip_id, "_w\\d+$"))) %>%
  group_by(id) %>%
  summarise(annual_system_loss = sum(cv_week, na.rm = TRUE), .groups = "drop")

# 3. View the Results
summary_stats <- total_system_welfare %>%
  summarise(
    Mean_Annual_Value = mean(annual_system_loss),
    Median_Annual_Value = median(annual_system_loss),
    Total_Sample_Value = sum(annual_system_loss)
  )

print(summary_stats)

```

# Coeffienct Table

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"
library(stargazer)
# Create a specific list for the 2018/2023 comparison
# We alternate them: [Obs 2018, 52wk 2018, Obs 2023, 52wk 2023]
comparison_list <- list(
    "Observed" = model_list_obs[["Year_2018"]], 
    "52-Week"  = model_list_52[["52wk 2018"]],
    "Observed" = model_list_obs[["Year_2023"]], 
    "52-Week"  = model_list_52[["52wk 2023"]]
)

# 2. Generate the Professionally Grouped Table
library(stargazer)

# Ensure your lists exist and are not empty before running this
stargazer(
  comparison_list,
  type = "html",
  out = "rum_comparison_results.html",
  title = "Table: Comparison of RUM Specifications (2018 vs. 2023)",
  
  # --- THE KEY ARGUMENT FOR SIDE-BY-SIDE SEs ---
  single.row = TRUE, 
  
  column.labels = c("2018", "2023"),
  column.separate = c(2, 2),
  model.names = TRUE,
  dep.var.labels.include = FALSE,
  model.numbers = FALSE,
  
  # Pulling Price (22) and Bunker (6) to the top
  order = c(22, 6),
  covariate.labels = c("Travel Cost (Price)", "Keanakolu Bunkhouse"),
  
  omit.stat = c("f", "ser", "lr", "rsq", "adj.rsq"),
  star.cutoffs = c(0.05, 0.01, 0.001),
  no.space = TRUE,
  notes = "Standard errors in parentheses next to coefficients. 52-Week model includes 'Stay-at-Home'."
)

```

# All Coefficent

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"

# Create a specific list for the 2018/2023 comparison

comparison_list <- list(
    "Observed" = model_list_obs[["Year_2018"]], 
    "52-Week"  = model_list_52[["52wk 2018"]],
    "Observed" = model_list_obs[["Year_2019"]], 
    "52-Week"  = model_list_52[["52wk 2019"]],
    "Observed" = model_list_obs[["Year_2020"]], 
    "52-Week"  = model_list_52[["52wk 2020"]],
    "Observed" = model_list_obs[["Year_2021"]], 
    "52-Week"  = model_list_52[["52wk 2021"]],
    "Observed" = model_list_obs[["Year_2022"]], 
    "52-Week"  = model_list_52[["52wk 2022"]],
    "Observed" = model_list_obs[["Year_2023"]], 
    "52-Week"  = model_list_52[["52wk 2023"]]
)

library(stargazer)
stargazer(
comparison_list,
type = "html",
out = "rum_comparison_results_appendix.html",
title = "Table: Comparison of RUM Specifications (2018 vs. 2023)",
column.labels = c("2018", "2019","2020","2021","2022","2023"),
column.separate = c(2, 2),
model.names = TRUE,
dep.var.labels.include = FALSE,
model.numbers = FALSE,
# Use 'order' to pull Price (22) and Bunker (6) to the top of the table
# This keeps the table concise and professional
order = c(23, 6),
# Rename only the ones we kept at the top
covariate.labels = c("Travel Cost (Price)", "Keanakolu Bunkhouse"),
omit.stat = c("f", "ser", "lr", "rsq", "adj.rsq"),
star.cutoffs = c(0.05, 0.01, 0.001),
no.space = TRUE,
notes = "Standard errors in parentheses. 52-Week model includes 'Stay-at-Home' as a weekly choice."
)

```

#### Merge

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"

library(dplyr)

final_comparison_master <- final_comparison %>%
  # Select and rename columns from the first dataset
  select(id, parkname, year, zip, income, 
         loss_rum_observed = annual_rum_loss, 
         loss_mdcev = mdcev_annual_loss) %>%
  

  full_join(
    final_comparison_52_person %>% 
      select(id, parkname, year, loss_rum_52wk = annual_rum_loss), 
    by = c("id", "parkname", "year")
  ) 

head(final_comparison_master)

```

## Welfare Table

```{r}
#| cache: true
library(dplyr)
library(tidyr)
library(stargazer)

# Calculate Mean and 95% Confidence Intervals
summary_wide <- final_comparison_master %>%
  filter(year %in% c(2018, 2023)) %>%
  group_by(parkname, year) %>%
  summarise(
    n = n(),
    mean_val = mean(loss_rum_observed, na.rm = TRUE),
    sd_val   = sd(loss_rum_observed, na.rm = TRUE),
    # Calculate Standard Error and 95% CI bounds
    se_val   = sd_val / sqrt(n),
    LCI      = mean_val - (1.96 * se_val),
    UCI      = mean_val + (1.96 * se_val),
    .groups = "drop"
  ) %>%
  select(parkname, year, mean_val, LCI, UCI) %>%
  pivot_wider(
    names_from = year, 
    values_from = c(mean_val, LCI, UCI),
    names_glue = "{year}_{.value}" 
  ) %>%
  # 2018 Stats then 2023 Stats
  select(parkname, 
         contains("2018"), 
         contains("2023")) %>%
  as.data.frame()



stargazer(summary_wide, 
          type = "html", 
          out = "Welfare_Comparison_Table.html",
          summary = FALSE, 
          rownames = FALSE,
          title = "Welfare Loss: Mean and 95% Confidence Intervals (2018 vs 2023)",
          digits = 2,
          # Groups the 3 columns under each year
          column.labels = c("Park Name", "2018", "2023"),
          column.separate = c(1, 3, 3), 
          covariate.labels = c("Park", 
                               "Mean", "Lower CI", "Upper CI", 
                               "Mean", "Lower CI", "Upper CI"))



# Calculate Mean and 95% Confidence Intervals
summary_wide <- final_comparison_master %>%
  filter(year %in% c(2018, 2023)) %>%
  group_by(parkname, year) %>%
  summarise(
    n = n(),
    mean_val = mean(loss_rum_52wk, na.rm = TRUE),
    sd_val   = sd(loss_rum_52wk, na.rm = TRUE),
    # Calculate Standard Error and 95% CI bounds
    se_val   = sd_val / sqrt(n),
    LCI      = mean_val - (1.96 * se_val),
    UCI      = mean_val + (1.96 * se_val),
    .groups = "drop"
  ) %>%
  select(parkname, year, mean_val, LCI, UCI) %>%
  pivot_wider(
    names_from = year, 
    values_from = c(mean_val, LCI, UCI),
    names_glue = "{year}_{.value}" 
  ) %>%
  # 2018 Stats then 2023 Stats
  select(parkname, 
         contains("2018"), 
         contains("2023")) %>%
  as.data.frame()



stargazer(summary_wide, 
          type = "html", 
          out = "Welfare_Comparison_Table_frequency.html",
          summary = FALSE, 
          rownames = FALSE,
          title = "Welfare Loss Frequency: Mean and 95% Confidence Intervals (2018 vs 2023)",
          digits = 2,
          # Groups the 3 columns under each year
          column.labels = c("Park Name", "2018", "2023"),
          column.separate = c(1, 3, 3), 
          covariate.labels = c("Park", 
                               "Mean", "Lower CI", "Upper CI", 
                               "Mean", "Lower CI", "Upper CI"))
```

```{r}
#| cache: true
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show the Code"


# Calculate Mean and 95% Confidence Intervals
summary_wide <- final_comparison_master %>%
  filter(year %in% c(2019, 2020)) %>%
  group_by(parkname, year) %>%
  summarise(
    n = n(),
    mean_val = mean(loss_rum_52wk, na.rm = TRUE),
    sd_val   = sd(loss_rum_52wk, na.rm = TRUE),
    # Calculate Standard Error and 95% CI bounds
    se_val   = sd_val / sqrt(n),
    LCI      = mean_val - (1.96 * se_val),
    UCI      = mean_val + (1.96 * se_val),
    .groups = "drop"
  ) %>%
  select(parkname, year, mean_val, LCI, UCI) %>%
  pivot_wider(
    names_from = year, 
    values_from = c(mean_val, LCI, UCI),
    names_glue = "{year}_{.value}" 
  ) %>%
  # 2018 Stats then 2023 Stats
  select(parkname, 
         contains("2019"), 
         contains("2020")) %>%
  as.data.frame()



stargazer(summary_wide, 
          type = "html", 
          out = "Welfare_Comparison_Table_frequency_1920.html",
          summary = FALSE, 
          rownames = FALSE,
          title = "Welfare Loss Frequency: Mean and 95% Confidence Intervals (2019 & 2020)",
          digits = 2,
          # Groups the 3 columns under each year
          column.labels = c("Park Name", "2019", "2020"),
          column.separate = c(1, 3, 3), 
          covariate.labels = c("Park", 
                               "Mean", "Lower CI", "Upper CI", 
                               "Mean", "Lower CI", "Upper CI"))


# ── Build summary for both models ─────────────────────────────────────────────
calc_summary <- function(data, loss_col, years) {
    data %>%
        filter(year %in% years, parkname != "Stay_Home") %>%
        group_by(parkname, year) %>%
        summarise(
            n        = n(),
            mean_val = mean(.data[[loss_col]],  na.rm = TRUE),
            sd_val   = sd(.data[[loss_col]],    na.rm = TRUE),
            se_val   = sd_val / sqrt(n),
            LCI      = mean_val - 1.96 * se_val,
            UCI      = mean_val + 1.96 * se_val,
            .groups  = "drop"
        ) %>%
        select(parkname, year, mean_val, LCI, UCI)
}

# ── Frequency RUM ─────────────────────────────────────────────────────────────
freq_wide <- calc_summary(final_comparison_master, "loss_rum_52wk", c(2021, 2022)) %>%
    pivot_wider(
        names_from  = year,
        values_from = c(mean_val, LCI, UCI),
        names_glue  = "{year}_freq_{.value}"
    )

# ── Observed RUM ──────────────────────────────────────────────────────────────
obs_wide <- calc_summary(final_comparison_master, "loss_rum_observed", c(2021, 2022)) %>%
    pivot_wider(
        names_from  = year,
        values_from = c(mean_val, LCI, UCI),
        names_glue  = "{year}_obs_{.value}"
    )

# ── Join both models ──────────────────────────────────────────────────────────
summary_wide <- freq_wide %>%
    left_join(obs_wide, by = "parkname") %>%
    select(
        parkname,
        # 2019 — Observed then Frequency
        `2021_obs_mean_val`, `2021_obs_LCI`, `2021_obs_UCI`,
        `2021_freq_mean_val`, `2021_freq_LCI`, `2021_freq_UCI`,
        # 2020 — Observed then Frequency
        `2022_obs_mean_val`, `2022_obs_LCI`, `2022_obs_UCI`,
        `2022_freq_mean_val`, `2022_freq_LCI`, `2022_freq_UCI`
    ) %>%
    arrange(parkname) %>%
    as.data.frame()

# ── Export ────────────────────────────────────────────────────────────────────
stargazer(
    summary_wide,
    type       = "html",
    out        = "Welfare_Comparison_Table_frequency_2122.html",
    summary    = FALSE,
    rownames   = FALSE,
    title      = "Welfare Loss: Observed and Frequency RUM — Mean and 95% CI (2021 & 2022)",
    digits     = 2,
    # 4 groups: 2020 Observed, 2020 Frequency, 2022 Observed, 2022 Frequency
    column.labels    = c("Park Name",
                         "2021 Observed", "2021 Frequency",
                         "2022 Observed", "2022 Frequency"),
    column.separate  = c(1, 3, 3, 3, 3),
    covariate.labels = c("Park",
                         "Mean", "Lower CI", "Upper CI",   # 2019 Observed
                         "Mean", "Lower CI", "Upper CI",   # 2019 Frequency
                         "Mean", "Lower CI", "Upper CI",   # 2020 Observed
                         "Mean", "Lower CI", "Upper CI"),  # 2020 Frequency
    notes       = "95% confidence intervals. Observed RUM conditions welfare on observed trips. Frequency RUM uses 52 weekly choice occasions.",
    notes.align = "l"
)

message("Saved to Welfare_Comparison_Table_frequency_2122.html")
```

#### Plot

So this model takes the park mean of the welfare per person and shows the mean per park across years. The stability shows that will including a frequency trip in RUM shows the highest variation, the RUM opt out is more concise and then the MDCEV is always in the middle.

```{r}
#| cache: true

# Summarize and Reshape the data
plot_data <- final_comparison_master %>%
  group_by(parkname, year) %>%
  filter(parkname!="Stay_Home")%>%
  summarise(
    ObservedRUM = mean(loss_rum_observed, na.rm = TRUE),
    FrequencyRUM = mean(loss_rum_52wk, na.rm = TRUE),
    MDCEV = mean(loss_mdcev, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(ObservedRUM, FrequencyRUM, MDCEV),
    names_to = "Model_Type",
    values_to = "Mean_Welfare_Loss"
  )


# Create the Boxplot 
p=ggplot(plot_data, aes(x = factor(year), y = Mean_Welfare_Loss, fill = Model_Type)) +
  # Classic boxplot with black outlines
  geom_boxplot(outlier.shape = 21, outlier.fill = "white", color = "black", alpha = 0.8) +
  
  # Labels
  labs(
    title = "Distribution of Mean Park Welfare Loss from Site Closures by Year",
    subtitle = "Comparing Observed RUM and MDCEV Model",
    x = "Year",
    y = "Average Welfare Loss Across sites ($)",
    fill = "Model Type"
  ) +
  
  # Colors
scale_fill_manual(values = c(
    "ObservedRUM"  = "#D9D9D9",   # Light Grey
    "MDCEV"        = "#252525"    # Dark Grey / Black
  )) +
  
  # Theme Customization
  theme_bw() + # Starting with theme_bw gives a clean white background
  theme(
    # Set all text to Times New Roman (serif)
    text = element_text(family = "serif", size =20),
    
    # Title and axis styling
    plot.title = element_text(face = "bold", size =26, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    
    # Remove gridlines for the "Classic/Blank" look
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    
    # Legend styling
    legend.position = "bottom",
    legend.background = element_blank(),
    
    # Classic full border (Using 'size' instead of 'linewidth' for compatibility)
    panel.border = element_rect(colour = "black", fill = NA, size = 1)
  )

p
ggsave("comparision_perperson_2models.png", plot = p,width=12, height=10, dpi=300)


```

## Per participant

```{r}
#| cache: true
full_year_clean=rbind(final_data_2018,final_data_2019,final_data_2020,final_data_2021,final_data_2022,final_data_2023)
# Identify participants by ensuring IDs match as characters
participants_only_welfare <- final_comparison_master %>%
    filter(parkname!="Stay_Home")%>%
  left_join(
    # Convert ID to character on the fly so the join works
    full_year_clean %>% 
      mutate(id = as.character(id)) %>% 
      select(id, parkname, year, VisitCount), 
    by = c("id", "parkname", "year")
  ) %>%
  # Filter for people who actually visited the park in question
  filter(VisitCount > 0)

# Summary for the "Participants Only" group
participants_summary_table <- participants_only_welfare %>%
  group_by(year, parkname) %>%
  summarise(
    avg_loss_obs_participant   = mean(loss_rum_observed, na.rm = TRUE),
    avg_loss_52wk_participant  = mean(loss_rum_52wk, na.rm = TRUE),
    avg_loss_mdcev_participant = mean(loss_mdcev, na.rm = TRUE),
    num_participants           = n(),
    .groups = "drop"
  )
```

#### Plot

```{r}
#| cache: true
# Reshape the Participants Summary for plotting
plot_data_participants <- participants_summary_table %>%
  # Select the three mean columns we created earlier
  select(year, parkname, 
         ObservedRUM = avg_loss_obs_participant, 
         FrequencyRUM = avg_loss_52wk_participant, 
         MDCEV = avg_loss_mdcev_participant) %>%
  pivot_longer(
    cols = c(ObservedRUM, FrequencyRUM, MDCEV),
    names_to = "Model_Type",
    values_to = "Mean_Welfare_Loss"
  )

#  Create the Grayscale Boxplot
participants=ggplot(plot_data_participants, aes(x = factor(year), y = Mean_Welfare_Loss, fill = Model_Type)) +
  # Standard boxplot with black outlines and white outliers
  geom_boxplot(outlier.shape = 21, outlier.fill = "white", color = "black", alpha = 1) +
  
  # Labels
  labs(
    title = "Mean Annual Welfare Loss: Active Participants Only",
    subtitle = "Calculated for individuals with Observed Visits (VisitCount > 0)",
    x = "Year",
    y = "Mean Annual Welfare Loss ($/User)",
    fill = "Model Type"
  ) +
  
  # Shades of Grey
  scale_fill_manual(values = c(
    "ObservedRUM"  = "#D9D9D9",   # Light Grey
    "FrequencyRUM" = "#737373",   # Medium Grey
    "MDCEV"        = "#252525"    # Dark Grey / Black
  )) +
  
  # Theme Customization (Blank Background + Times New Roman)
  theme_bw() + 
  theme(
    text = element_text(family = "serif", size = 20),
    plot.title = element_text(face = "bold", size = 26, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    
    # Remove all gridlines for a completely blank white background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    
    # Legend styling
    legend.position = "bottom",
    legend.background = element_blank(),
    
    # Classic full black border around the plot (using size for compatibility)
    panel.border = element_rect(colour = "black", fill = NA, size = 1)
  )

ggsave("comparision_perparticipate.png",plot=participants, width=12, height=10, dpi=300)

participants

```

### EJ

Now that we have the per person welfare across individuals we can link the zip to a location with a higher EJ score and see how this spread spills over to EJ

```{r}
#| cache: true
# Hud data
library(readxl)
ZIP_TRACT_122025 <- read_excel("data/ZIP_TRACT_122025.xlsx")

# only hawaii

dishawaii <- readRDS("data/dishawaii.rds")

# clean tract
dishawaii <- dishawaii %>%
    mutate(TRACT = str_pad(as.character(ID), 11, "left", "0"))

# This taking the tact by zip code filtering hawaii  and making sure tract and zip in write order
# THIS is where you need to make changes
# It should be in an earlier chunk in your script

dishawaii_tract <- dishawaii %>%
  rowwise() %>%
  mutate(
    # ── CHANGE 1: threshold from >= 80 to >= 90 ───────────────────────────
    env_burden = any(c_across(c(
      P_PM25, P_OZONE, P_DSLPM, P_RSEI_AIR, P_PTRAF,
      P_LDPNT, P_PNPL, P_PRMP, P_PTSDF, P_UST, P_PWDIS,
      P_NO2, P_DWATER,
      P_D2_PM25, P_D5_PM25, P_D2_OZONE, P_D5_OZONE,
      P_D2_DSLPM, P_D5_DSLPM, P_D2_RSEI_AIR, P_D5_RSEI_AIR,
      P_D2_PTRAF, P_D5_PTRAF, P_D2_LDPNT, P_D5_LDPNT,
      P_D2_PNPL, P_D5_PNPL, P_D2_PRMP, P_D5_PRMP,
      P_D2_PTSDF, P_D5_PTSDF, P_D2_UST, P_D5_UST,
      P_D2_PWDIS, P_D5_PWDIS, P_D2_NO2, P_D5_NO2,
      P_D2_DWATER, P_D5_DWATER
    )) >= 90,   # <-- CHANGE THIS NUMBER (80 or 90)
    na.rm = TRUE),

    soc_vulnerable = any(c_across(c(
      P_PEOPCOLORPCT, P_LOWINCPCT, P_UNEMPPCT, P_DISABILITYPCT,
      P_LINGISOPCT, P_LESSHSPCT, P_UNDER5PCT, P_OVER64PCT, P_LIFEEXPPCT
    )) >= 90,   # <-- CHANGE THIS NUMBER (80 or 90)
    na.rm = TRUE),

    # ── CHANGE 2: AND vs OR logic ─────────────────────────────────────────
    DAC_flag = as.integer(env_burden | soc_vulnerable)
    #                               ^ change & to | for OR logic
  ) %>%
  ungroup()


hud_crosswalk <- ZIP_TRACT_122025 %>%
    rename_with(toupper) %>% 
    mutate(
        # Ensure standard 11-digit Tract and 5-digit ZIP
        TRACT = str_pad(as.character(TRACT), 11, "left", "0"),
        ZIP   = str_pad(as.character(ZIP), 5, "left", "0")
    ) %>%
    # Filter for Hawaii only: 
    # Either by ZIP prefix (967/968) OR Census Tract State prefix (15)
    filter(str_starts(TRACT, "15"))


#Define thursholds into the 90s
dishawaii_zip <- hud_crosswalk %>%
  left_join(dishawaii_tract, by = "TRACT") %>%
  group_by(ZIP) %>%
  summarise(
    
    # ── Binary population share flags ────────────────────────────────────────
    pct_pop_in_DAC  = sum(TOT_RATIO[DAC_flag == 1],          na.rm = TRUE),
    pct_pop_in_Envi = sum(TOT_RATIO[env_burden == TRUE],     na.rm = TRUE),
    pct_pop_in_Soci = sum(TOT_RATIO[soc_vulnerable == TRUE], na.rm = TRUE),
    
    # ── Environmental composite ───────────────────────────────────────────────
    # Use P_ columns — state percentiles, all on 0-100 scale
    # Excludes PM25 and OZONE (limited Hawaii data)
    env_composite = weighted.mean(
      rowMeans(cbind(
        P_DSLPM,    P_RSEI_AIR, P_PTRAF,  P_LDPNT,
        P_PNPL,     P_PRMP,     P_PTSDF,  P_UST,
        P_PWDIS,    P_NO2,      P_DWATER
      ), na.rm = TRUE),
      TOT_RATIO, na.rm = TRUE
    ),
    
    # ── Social vulnerability composite ────────────────────────────────────────
    # All 9 social P_ indicators — state percentiles
    soci_composite = weighted.mean(
      rowMeans(cbind(
        P_PEOPCOLORPCT, P_LOWINCPCT,    P_UNEMPPCT,
        P_DISABILITYPCT, P_LINGISOPCT,  P_LESSHSPCT,
        P_UNDER5PCT,    P_OVER64PCT,    P_LIFEEXPPCT
      ), na.rm = TRUE),
      TOT_RATIO, na.rm = TRUE
    ),
    
    # ── Also store the index scores EJScreen already computed ─────────────────
    # DEMOGIDX_2 is EJScreen's own demographic index — useful as a check
    demog_index = weighted.mean(P_DEMOGIDX_2, TOT_RATIO, na.rm = TRUE),
    
    total_ratio = sum(TOT_RATIO, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    
    # ── Binary flags ──────────────────────────────────────────────────────────
    is_disadvantaged_50      = ifelse(pct_pop_in_DAC  >= 1, 1, 0),
    is_disadvantaged_50_envi = ifelse(pct_pop_in_Envi >= 1, 1, 0),
    is_disadvantaged_50_soci = ifelse(pct_pop_in_Soci >= 1, 1, 0),
    
    # ── Environmental percentile blocks ───────────────────────────────────────
    # High env_pctile
    env_pctile = ntile(env_composite, 100),
    env_block  = cut(env_pctile,
      breaks = c(0, 20, 40, 60, 80, 90, 95, 100),
      labels = c(
        "0-20th\n(Lowest)",  "20-40th", "40-60th",
        "60-80th", "80-90th", "90-95th", "95-100th\n(Highest)"
      ),
      include.lowest = TRUE
    ),
    
    # ── Social vulnerability percentile blocks ────────────────────────────────
    # High soci_pctile = high vulnerability = most vulnerable
    soci_pctile = ntile(soci_composite, 100),
    soci_block  = cut(soci_pctile,
      breaks = c(0, 20, 40, 60, 80, 90, 95, 100),
      labels = c(
        "0-20th\n(Lowest)",  "20-40th", "40-60th",
        "60-80th", "80-90th", "90-95th", "95-100th\n(Highest)"
      ),
      include.lowest = TRUE
    )
  )

# ── Verification checks ───────────────────────────────────────────────────────

# 1. Composite scores should increase with block
dishawaii_zip %>%
  group_by(env_block) %>%
  summarise(
    mean_env  = round(mean(env_composite,  na.rm = TRUE), 1),
    mean_flag = round(mean(pct_pop_in_Envi, na.rm = TRUE), 2),
    n         = n()
  ) %>% arrange(env_block)

dishawaii_zip %>%
  group_by(soci_block) %>%
  summarise(
    mean_soci = round(mean(soci_composite,  na.rm = TRUE), 1),
    mean_flag = round(mean(pct_pop_in_Soci, na.rm = TRUE), 2),
    n         = n()
  ) %>% arrange(soci_block)

# 2. Check coverage
dishawaii_zip %>%
  summarise(
    total_zips      = n(),
    env_complete    = sum(!is.na(env_composite)),
    soci_complete   = sum(!is.na(soci_composite)),
    total_ratio_avg = round(mean(total_ratio), 2)  # should be close to 1.0
  )

# ── Join to welfare data ──────────────────────────────────────────────────────
final_comparison_master1 <- final_comparison_master %>%
  left_join(
    dishawaii_zip %>%
      mutate(ZIP = as.numeric(ZIP)) %>%   # convert ZIP to numeric here
      select(ZIP,
             env_composite,  env_pctile,  env_block,
             soci_composite, soci_pctile, soci_block,
             demog_index,
             pct_pop_in_DAC, pct_pop_in_Envi, pct_pop_in_Soci,
             is_disadvantaged_50, is_disadvantaged_50_envi, is_disadvantaged_50_soci),
    by = c("zip" = "ZIP")
  )


```

# Income

```{r}
#| cache: true
final_comparison_master1 <- final_comparison_master1 %>%
  mutate(
    income_pctile = 101 - ntile(income, 100),
    
    income_block = cut(income_pctile,
      breaks = c(0, 20, 40, 60, 80, 90, 95, 100),
      labels = c(
        "0-20th\n(Lowest)",  "20-40th",  "40-60th",
        "60-80th", "80-90th", "90-95th", "95-100th\n(Highest)"
      ),
      include.lowest = TRUE
    )
  )

avg_data <- final_comparison_master1 %>%
  filter(year %in% c(2018, 2023), parkname !="Stay_Home") %>%
  pivot_longer(
    cols      = c(loss_mdcev, loss_rum_observed,loss_rum_52wk),
    names_to  = "model",
    values_to = "loss"
  ) %>%
  mutate(
    model = recode(model,
      "loss_mdcev"        = "MDCEV",
      "loss_rum_observed" = "Observed RUM",
      "loss_rum_52wk" = "Frequency RUM"
    )
  ) %>%
  group_by(income_block, model, year) %>%
  summarise(
    mean_loss = mean(loss,           na.rm = TRUE),
    se        = sd(loss, na.rm = TRUE) / sqrt(n()),
    p25       = quantile(loss, 0.25, na.rm = TRUE),
    p75       = quantile(loss, 0.75, na.rm = TRUE),
    n         = n(),
    .groups   = "drop"
  ) %>%
  mutate(
    ci_low  = mean_loss - 1.96 * se,
    ci_high = mean_loss + 1.96 * se,
    year    = factor(year, labels = c("2018", "2023"))
  )


avg_data %>% distinct(income_block) %>% arrange(income_block)

# ── Shade the 90th+ blocks  ──────────────────
vulnerable_positions <- avg_data %>%
  filter(grepl("90-95th|95-100th", as.character(income_block))) %>%
  mutate(block_num = as.numeric(income_block)) %>%
  distinct(block_num) %>%
  pull(block_num)

shade_start <- min(vulnerable_positions) - 0.5
shade_end   <- max(vulnerable_positions) + 0.5

# ── Plot ──────────────────────────────────────────────────────────────────────
# ── Step 1: Build each indicator separately with identical structure ──────────

# Income — already built, just add indicator column
income_avg <- final_comparison_master1 %>%
    filter(year %in% c(2018, 2023), !is.na(income_block), parkname !="Stay_Home") %>%
    pivot_longer(
        cols      = c(loss_mdcev, loss_rum_observed,loss_rum_52wk ),
        names_to  = "model",
        values_to = "loss"
    ) %>%
    mutate(
        model        = recode(model,
                              "loss_mdcev"        = "MDCEV",
                              "loss_rum_observed" = "Observed RUM",
      "loss_rum_52wk" = "Frequency RUM"),
        pctile_block = income_block,
        indicator    = "Income\n(Least → Most Vulnerable)"
    )

# Environmental burden — high pctile = most vulnerable = already RIGHT
env_avg <- final_comparison_master1 %>%
    filter(year %in% c(2018, 2023), !is.na(env_block), parkname !="Stay_Home") %>%
    pivot_longer(
        cols      = c(loss_mdcev, loss_rum_observed,loss_rum_52wk),
        names_to  = "model",
        values_to = "loss"
    ) %>%
    mutate(
        model        = recode(model,
                              "loss_mdcev"        = "MDCEV",
                              "loss_rum_observed" = "Observed RUM",
      "loss_rum_52wk" = "Frequency RUM"),
        pctile_block = env_block,
        indicator    = "Environmental Burden\n(Least → Most Vulnerable)"
    )

# Social vulnerability — high pctile = most vulnerable = already RIGHT
soci_avg <- final_comparison_master1 %>%
    filter(year %in% c(2018, 2023), !is.na(soci_block), parkname !="Stay_Home") %>%
    pivot_longer(
        cols      = c(loss_mdcev, loss_rum_observed,loss_rum_52wk),
        names_to  = "model",
        values_to = "loss"
    ) %>%
    mutate(
        model        = recode(model,
                              "loss_mdcev"        = "MDCEV",
                              "loss_rum_observed" = "Observed RUM",
      "loss_rum_52wk" = "Frequency RUM"),
        pctile_block = soci_block,
        indicator    = "Social Vulnerability\n(Least → Most Vulnerable)"
    )

# ── Step 2: Combine and summarise ─────────────────────────────────────────────
all_avg <- bind_rows(income_avg, env_avg, soci_avg) %>%
    mutate(
        year = factor(year, labels = c("2018", "2023")),
        indicator = factor(indicator, levels = c(
            "Income\n(Least → Most Vulnerable)",
            "Social Vulnerability\n(Least → Most Vulnerable)",
            "Environmental Burden\n(Least → Most Vulnerable)"
        ))
    ) %>%
    group_by(indicator, pctile_block, model, year) %>%
    summarise(
        mean_loss = mean(loss,           na.rm = TRUE),
        se        = sd(loss, na.rm = TRUE) / sqrt(n()),
        p25       = quantile(loss, 0.25, na.rm = TRUE),
        p75       = quantile(loss, 0.75, na.rm = TRUE),
        n         = n(),
        .groups   = "drop"
    ) %>%
    mutate(
        ci_low  = mean_loss - 1.96 * se,
        ci_high = mean_loss + 1.96 * se
    )

# ── Step 3: Shading positions — 90th+ blocks are RIGHT in all three ───────────
vulnerable_positions <- all_avg %>%
    filter(grepl("90-95th|95-100th", as.character(pctile_block))) %>%
    mutate(block_num = as.numeric(pctile_block)) %>%
    distinct(block_num) %>%
    pull(block_num)

shade_start <- min(vulnerable_positions) - 0.5
shade_end   <- max(vulnerable_positions) + 0.5
```

```{r}
#| cache: true
# ── Step 4: Plot ──────────────────────────────────────────────────────────────
ggplot(all_avg,
       aes(x     = pctile_block,
           y     = mean_loss,
           color = model,
           group = model)) +
  
  geom_rect(
    data = tibble(xmin = shade_start, xmax = shade_end),
    aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
    fill        = "gray90",
    alpha       = 0.5,
    inherit.aes = FALSE
  ) +
  
  geom_errorbar(
    aes(ymin = ci_low, ymax = ci_high),
    width     = 0.2,
    linewidth = 0.4,
    position  = position_dodge(0.4)   # slightly wider dodge for 3 models
  ) +
  
  geom_line(linewidth = 0.8,  position = position_dodge(0.4)) +
  geom_point(size = 2.5,      position = position_dodge(0.4)) +
  
  geom_hline(
    yintercept = 0, linetype = "dashed",
    color = "gray40", linewidth = 0.4
  ) +
  
  geom_text(
    data = tibble(
      pctile_block = factor("90-95th",
                            levels = levels(all_avg$pctile_block)),
      mean_loss    = -2,
      year         = factor("2018"),
      indicator    = factor("Income\n(Least → Most Vulnerable)",
                            levels = levels(all_avg$indicator))
    ),
    aes(x = pctile_block, y = mean_loss),
    label       = "90th+\nVulnerable",
    inherit.aes = FALSE,
    size        = 3.5,
    color       = "gray40",
    hjust       = 0.5,
    family      = "Times New Roman"
  ) +
  
  facet_grid(indicator ~ year) +
  
  # ── Three model grayscale ─────────────────────────────────────────────────
  scale_color_manual(
    values = c(
      "MDCEV"         = "gray10",
      "Observed RUM"  = "gray55",
      "Frequency RUM" = "gray80"    # lightest gray for third model
    ),
    name = "Model"
  ) +
  scale_linetype_manual(
    values = c(
      "MDCEV"         = "solid",
      "Observed RUM"  = "solid",
      "Frequency RUM" = "dashed"    # dashed helps distinguish in grayscale
    ),
    name = "Model"
  ) +
  aes(linetype = model) +           # add linetype aesthetic
  
  scale_y_continuous(
    labels = scales::dollar_format(prefix = "$"),
    name   = "Mean Welfare Loss"
  ) +
  
  labs(
    x     = "Percentile Block",
    title = "Mean Welfare Loss Across Distributions"
  ) +
  
  theme_minimal(base_family = "Times New Roman") +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text.x   = element_text(size = 16, angle = 20,
                                 hjust = 1, lineheight = 0.8),
    axis.text.y   = element_text(size = 20),
    strip.text.x  = element_text(face = "bold", size = 16),
    strip.text.y  = element_text(face = "bold", size = 20,
                                 angle = 0, hjust = 0),
    legend.position = "bottom",
    legend.title  = element_text(face = "bold", size = 20),
    legend.text   = element_text(size = 20),
    plot.title    = element_text(face = "bold", size = 20),
    panel.spacing = unit(1.2, "lines")
  )
ggsave("social_vulnerabilities.png", width=12, height=10, dpi=300)

```

```{r}
#| cache: true
# ── Rebuild all_avg_appendix with three models ────────────────────────────────

income_avg_all <- final_comparison_master1 %>%
  filter(!is.na(income_block), parkname!="Stay_Home") %>%
  pivot_longer(
    cols      = c(loss_mdcev, loss_rum_observed, loss_rum_52wk),  # <-- added
    names_to  = "model",
    values_to = "loss"
  ) %>%
  mutate(
    model = recode(model,
      "loss_mdcev"        = "MDCEV",
      "loss_rum_observed" = "Observed RUM",
      "loss_rum_52wk"     = "Frequency RUM"   # <-- added
    ),
    pctile_block = income_block,
    indicator    = "Income\n(Least → Most Vulnerable)"
  )

env_avg_all <- final_comparison_master1 %>%
  filter(!is.na(env_block), parkname!="Stay_Home") %>%
  pivot_longer(
    cols      = c(loss_mdcev, loss_rum_observed, loss_rum_52wk),
    names_to  = "model",
    values_to = "loss"
  ) %>%
  mutate(
    model = recode(model,
      "loss_mdcev"        = "MDCEV",
      "loss_rum_observed" = "Observed RUM",
      "loss_rum_52wk"     = "Frequency RUM"
    ),
    pctile_block = env_block,
    indicator    = "Environmental Burden\n(Least → Most Vulnerable)"
  )

soci_avg_all <- final_comparison_master1 %>%
  filter(!is.na(soci_block), parkname!="Stay_Home") %>%
  pivot_longer(
    cols      = c(loss_mdcev, loss_rum_observed, loss_rum_52wk),
    names_to  = "model",
    values_to = "loss"
  ) %>%
  mutate(
    model = recode(model,
      "loss_mdcev"        = "MDCEV",
      "loss_rum_observed" = "Observed RUM",
      "loss_rum_52wk"     = "Frequency RUM"
    ),
    pctile_block = soci_block,
    indicator    = "Social Vulnerability\n(Least → Most Vulnerable)"
  )

all_avg_appendix <- bind_rows(income_avg_all, env_avg_all, soci_avg_all) %>%
  mutate(
    year      = factor(year),
    model     = factor(model, levels = c("MDCEV", "Observed RUM", "Frequency RUM")),
    indicator = factor(indicator, levels = c(
      "Income\n(Least → Most Vulnerable)",
      "Social Vulnerability\n(Least → Most Vulnerable)",
      "Environmental Burden\n(Least → Most Vulnerable)"
    ))
  ) %>%
  group_by(indicator, pctile_block, model, year) %>%
  summarise(
    mean_loss = mean(loss,           na.rm = TRUE),
    se        = sd(loss, na.rm = TRUE) / sqrt(n()),
    p25       = quantile(loss, 0.25, na.rm = TRUE),
    p75       = quantile(loss, 0.75, na.rm = TRUE),
    n         = n(),
    .groups   = "drop"
  ) %>%
  mutate(
    ci_low  = mean_loss - 1.96 * se,
    ci_high = mean_loss + 1.96 * se
  )

# ── Updated helper function with three models ─────────────────────────────────
make_appendix_plot <- function(data, years, title_suffix) {
  
  plot_data <- data %>% filter(year %in% years)
  
  vulnerable_positions <- plot_data %>%
    filter(grepl("90-95th|95-100th", as.character(pctile_block))) %>%
    mutate(block_num = as.numeric(pctile_block)) %>%
    distinct(block_num) %>%
    pull(block_num)
  
  shade_start <- min(vulnerable_positions) - 0.5
  shade_end   <- max(vulnerable_positions) + 0.5
  
  ggplot(plot_data,
         aes(x        = pctile_block,
             y        = mean_loss,
             color    = model,
             linetype = model,        # <-- added for readability
             group    = model)) +
    
    geom_rect(
      data = tibble(xmin = shade_start, xmax = shade_end),
      aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
      fill        = "gray90",
      alpha       = 0.5,
      inherit.aes = FALSE
    ) +
    
    geom_errorbar(
      aes(ymin = ci_low, ymax = ci_high),
      width     = 0.2,
      linewidth = 0.35,
      position  = position_dodge(0.4)   # slightly wider for 3 models
    ) +
    
    geom_line(linewidth = 0.7,  position = position_dodge(0.4)) +
    geom_point(size = 1.8,      position = position_dodge(0.4)) +
    
    geom_hline(
      yintercept = 0, linetype = "dashed",
      color = "gray40", linewidth = 0.4
    ) +
    
    facet_grid(indicator ~ year) +
    
    # ── Three model color and linetype scales ─────────────────────────────────
    scale_color_manual(
      values = c(
        "MDCEV"         = "gray10",
        "Observed RUM"  = "gray55",
        "Frequency RUM" = "gray75"    # <-- added
      ),
      name = "Model"
    ) +
    scale_linetype_manual(
      values = c(
        "MDCEV"         = "solid",
        "Observed RUM"  = "solid",
        "Frequency RUM" = "dashed"    # <-- dashed to distinguish from Observed RUM
      ),
      name = "Model"
    ) +
    scale_fill_manual(
      values = c(
        "MDCEV"         = "gray10",
        "Observed RUM"  = "gray55",
        "Frequency RUM" = "gray75"
      ),
      guide  = "none"
    ) +
    scale_y_continuous(
      labels = scales::dollar_format(prefix = "$"),
      name   = "Mean Welfare Loss"
    ) +
    
    labs(
      x     = "Percentile Block",
      title = paste("Mean Welfare Loss Across Distributions —", title_suffix)
    ) +
    
    theme_minimal(base_family = "Times New Roman") +
    theme(
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.x   = element_text(size = 12, angle = 30,
                                   hjust = 1, lineheight = 0.8),
      axis.text.y   = element_text(size = 20),
      strip.text.x  = element_text(face = "bold", size = 12),
      strip.text.y  = element_text(face = "bold", size = 14,
                                   angle = 0, hjust = 0),
      legend.position  = "bottom",
      legend.title     = element_text(face = "bold", size = 20),
      legend.text      = element_text(size = 20),
      plot.title       = element_text(face = "bold", size = 20),
      panel.spacing    = unit(0.8, "lines")
    )
}

# ── Build and save ────────────────────────────────────────────────────────────
appendix_a <- make_appendix_plot(
  all_avg_appendix,
  years        = c(2018, 2019, 2020),
  title_suffix = "2018–2020"
)

appendix_b <- make_appendix_plot(
  all_avg_appendix,
  years        = c(2021, 2022, 2023),
  title_suffix = "2021–2023"
)


ggsave(
  "appendix_A_welfare_2018_2020.png",
  plot   = appendix_a,
  width  = 14,
  height = 12,
  dpi    = 300,
  bg     = "white"
)

ggsave(
  "appendix_B_welfare_2021_2023.png",
  plot   = appendix_b,
  width  = 14,
  height = 12,
  dpi    = 300,
  bg     = "white"
)
print(appendix_a)

```

# Spread Across Models & DCA

The spread here show the value of park based on whether or not the person comes from a disadvantage community.

The MDCEV is fairly stable compared to the RUM and the spread is larger whe nconsidering each trip occasion.

For DCA are consistently less in Observed RUM and MDCEV but is greater in Frequency 52.

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)

# 1. Prepare data: Group by Park, Year, and EJ Status
plot_data_three_panel <- final_comparison_master1 %>%
  filter(parkname!="Stay_Home") %>%
  group_by(year, parkname, is_disadvantaged_50) %>%
  summarise(
    Observed_RUM = mean(loss_rum_observed, na.rm = TRUE),
    Frequency_RUM = mean(loss_rum_52wk, na.rm = TRUE),
    MDCEV = mean(loss_mdcev, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Pivot to long format
  pivot_longer(
    cols = c(Observed_RUM, Frequency_RUM, MDCEV),
    names_to = "Model_Type",
    values_to = "Welfare_Loss"
  ) %>%
  # Clean up factor labels for the legend and facets
  mutate(
    Model_Type = factor(Model_Type, 
                        levels = c("Observed_RUM", "Frequency_RUM", "MDCEV"),
                        labels = c("Observed RUM", "Frequency RUM (52-Wk)", "MDCEV")),
    EJ_Group = factor(is_disadvantaged_50, 
                      levels = c(0, 1), 
                      labels = c("Non-Disadvantaged", "Disadvantaged (EJ)"))
  )

# 2. Create the Three-Panel Plot
ggplot(plot_data_three_panel, aes(x = factor(year), y = Welfare_Loss, fill = EJ_Group)) +
  # Boxplot with black outlines
  geom_boxplot(outlier.shape = 21, outlier.size = 1, color = "black", alpha = 0.8) +
  
  # CREATE THREE PANELS (One for each Model)
  facet_wrap(~ Model_Type, scales = "free_y") + 
  
  # Grayscale for the EJ Comparison
  scale_fill_manual(values = c(
    "Non-Disadvantaged" = "#F0F0F0", # Very Light Grey
    "Disadvantaged (EJ)" = "#636363"  # Dark Grey
  )) +
  
  # Labels
  labs(
    title = "Annual Welfare Loss by Model Type and EJ Status",
    subtitle = "Comparing Disadvantaged vs. Non-Disadvantaged ZIP Codes across Hawaii Parks",
    x = "Year",
    y = "Average Annual Welfare Loss ($/Park)",
    fill = "Community Status"
  ) +
  
  # Academic Theme
  theme_bw() +
  theme(
    text = element_text(family = "serif", size = 16),
    plot.title = element_text(face = "bold", size = 20, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 16),
    
    # Grid and Panel styling
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(face = "bold", size = 16),
    
    # Legend at bottom
    legend.position = "bottom",
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8)
  )

ggsave("comparision_EJstatus.png", width=12, height=10, dpi=300)

```

```{r}

# 1. Prepare data: Group by Park, Year, and EJ Status
plot_data_three_panel <- final_comparison_master1 %>%
  filter(parkname!="Stay_Home") %>%
  group_by(year, parkname, is_disadvantaged_50_envi) %>%
  summarise(
    Observed_RUM = mean(loss_rum_observed, na.rm = TRUE),
    Frequency_RUM = mean(loss_rum_52wk, na.rm = TRUE),
    MDCEV = mean(loss_mdcev, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Pivot to long format
  pivot_longer(
    cols = c(Observed_RUM, Frequency_RUM, MDCEV),
    names_to = "Model_Type",
    values_to = "Welfare_Loss"
  ) %>%
  # Clean up factor labels for the legend and facets
  mutate(
    Model_Type = factor(Model_Type, 
                        levels = c("Observed_RUM", "Frequency_RUM", "MDCEV"),
                        labels = c("Observed RUM", "Frequency RUM (52-Wk)", "MDCEV")),
    EJ_Group = factor(is_disadvantaged_50_envi, 
                      levels = c(0, 1), 
                      labels = c("Non-Disadvantaged", "Disadvantaged (EJ)"))
  )

# 2. Create the Three-Panel Plot
ggplot(plot_data_three_panel, aes(x = factor(year), y = Welfare_Loss, fill = EJ_Group)) +
  # Boxplot with black outlines
  geom_boxplot(outlier.shape = 21, outlier.size = 1, color = "black", alpha = 0.8) +
  
  # CREATE THREE PANELS (One for each Model)
  facet_wrap(~ Model_Type, scales = "free_y") + 
  
  # Grayscale for the EJ Comparison
  scale_fill_manual(values = c(
    "Non-Disadvantaged" = "#F0F0F0", # Very Light Grey
    "Disadvantaged (EJ)" = "#636363"  # Dark Grey
  )) +
  
  # Labels
  labs(
    title = "Annual Welfare Loss by Model Type and Environmental Burdens Status",
    subtitle = "Comparing Disadvantaged vs. Non-Disadvantaged ZIP Codes across Hawaii Parks",
    x = "Year",
    y = "Average Annual Welfare Loss ($/Park)",
    fill = "Community Status"
  ) +
  
  # Academic Theme
  theme_bw() +
  theme(
    text = element_text(family = "serif", size = 12),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    
    # Grid and Panel styling
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(face = "bold", size = 11),
    
    # Legend at bottom
    legend.position = "bottom",
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8)
  )

ggsave("comparision_envir_status.png", width=12, height=10, dpi=300)

# 1. Prepare data: Group by Park, Year, and EJ Status
plot_data_three_panel <- final_comparison_master1 %>%
  filter(parkname!="Stay_Home") %>%
  group_by(year, parkname, is_disadvantaged_50_envi) %>%
  summarise(
    Observed_RUM = mean(loss_rum_observed, na.rm = TRUE),
    MDCEV = mean(loss_mdcev, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Pivot to long format
  pivot_longer(
    cols = c(Observed_RUM, MDCEV),
    names_to = "Model_Type",
    values_to = "Welfare_Loss"
  ) %>%
  # Clean up factor labels for the legend and facets
  mutate(
    Model_Type = factor(Model_Type, 
                        levels = c("Observed_RUM",  "MDCEV"),
                        labels = c("Observed RUM", "MDCEV")),
    EJ_Group = factor(is_disadvantaged_50_envi, 
                      levels = c(0, 1), 
                      labels = c("Non-Disadvantaged", "Disadvantaged"))
  )

# 2. Create the Three-Panel Plot
ggplot(plot_data_three_panel, aes(x = factor(year), y = Welfare_Loss, fill = EJ_Group)) +
  # Boxplot with black outlines
  geom_boxplot(outlier.shape = 21, outlier.size = 1, color = "black", alpha = 0.8) +
  
  # CREATE THREE PANELS (One for each Model)
  facet_wrap(~ Model_Type, scales = "free_y") + 
  
  # Grayscale for the EJ Comparison
  scale_fill_manual(values = c(
    "Non-Disadvantaged" = "#F0F0F0", # Very Light Grey
    "Disadvantaged" = "#636363"  # Dark Grey
  )) +
  
  # Labels
  labs(
    title = "Annual Welfare Loss by Model Type and Environmental Burdens Status",
    subtitle = "Comparing Disadvantaged vs. Non-Disadvantaged ZIP Codes across Hawaii Parks",
    x = "Year",
    y = "Average Annual Welfare Loss ($/Park)",
    fill = "Community Status"
  ) +
  
  # Academic Theme
  theme_bw() +
  theme(
    text = element_text(family = "serif", size = 20),
    plot.title = element_text(face = "bold", size =20, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 20),
    
    # Grid and Panel styling
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(face = "bold", size = 20),
    
    # Legend at bottom
    legend.position = "bottom",
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8)
  )

ggsave("comparision_envir_observed.png", width=12, height=10, dpi=300)


```

```{r}

# 1. Prepare data: Group by Park, Year, and EJ Status
plot_data_three_panel <- final_comparison_master1 %>%
  filter(parkname!="Stay_Home") %>%
  group_by(year, parkname, is_disadvantaged_50_soci) %>%
  summarise(
    Observed_RUM = mean(loss_rum_observed, na.rm = TRUE),
    Frequency_RUM = mean(loss_rum_52wk, na.rm = TRUE),
    MDCEV = mean(loss_mdcev, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Pivot to long format
  pivot_longer(
    cols = c(Observed_RUM, Frequency_RUM, MDCEV),
    names_to = "Model_Type",
    values_to = "Welfare_Loss"
  ) %>%
  # Clean up factor labels for the legend and facets
  mutate(
    Model_Type = factor(Model_Type, 
                        levels = c("Observed_RUM", "Frequency_RUM", "MDCEV"),
                        labels = c("Observed RUM", "Frequency RUM (52-Wk)", "MDCEV")),
    EJ_Group = factor(is_disadvantaged_50_soci, 
                      levels = c(0, 1), 
                      labels = c("Non-Disadvantaged", "Disadvantaged (EJ)"))
  )

# 2. Create the Three-Panel Plot
ggplot(plot_data_three_panel, aes(x = factor(year), y = Welfare_Loss, fill = EJ_Group)) +
  # Boxplot with black outlines
  geom_boxplot(outlier.shape = 21, outlier.size = 1, color = "black", alpha = 0.8) +
  
  # CREATE THREE PANELS (One for each Model)
  facet_wrap(~ Model_Type, scales = "free_y") + 
  
  # Grayscale for the EJ Comparison
  scale_fill_manual(values = c(
    "Non-Disadvantaged" = "#F0F0F0", # Very Light Grey
    "Disadvantaged (EJ)" = "#636363"  # Dark Grey
  )) +
  
  # Labels
  labs(
    title = "Annual Welfare Loss by Model Type and Social Vulnerabilities Status",
    subtitle = "Comparing Disadvantaged vs. Non-Disadvantaged ZIP Codes across Hawaii Parks",
    x = "Year",
    y = "Average Annual Welfare Loss ($/Park)",
    fill = "Community Status"
  ) +
  
  # Academic Theme
  theme_bw() +
  theme(
    text = element_text(family = "serif", size = 12),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    
    # Grid and Panel styling
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(face = "bold", size = 11),
    
    # Legend at bottom
    legend.position = "bottom",
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8)
  )
ggsave("comparision_social_vulnerabilities.png", width=12, height=10, dpi=300)



# 1. Prepare data: Group by Park, Year, and EJ Status
plot_data_three_panel <- final_comparison_master1 %>%
  filter(parkname!="Stay_Home") %>%
  group_by(year, parkname, is_disadvantaged_50_soci) %>%
  summarise(
    Observed_RUM = mean(loss_rum_observed, na.rm = TRUE),
    MDCEV = mean(loss_mdcev, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Pivot to long format
  pivot_longer(
    cols = c(Observed_RUM, MDCEV),
    names_to = "Model_Type",
    values_to = "Welfare_Loss"
  ) %>%
  # Clean up factor labels for the legend and facets
  mutate(
    Model_Type = factor(Model_Type, 
                        levels = c("Observed_RUM", "MDCEV"),
                        labels = c("Observed RUM", "MDCEV")),
    EJ_Group = factor(is_disadvantaged_50_soci, 
                      levels = c(0, 1), 
                      labels = c("Non-Disadvantaged", "Disadvantaged"))
  )

# 2. Create the Three-Panel Plot
ggplot(plot_data_three_panel, aes(x = factor(year), y = Welfare_Loss, fill = EJ_Group)) +
  # Boxplot with black outlines
  geom_boxplot(outlier.shape = 21, outlier.size = 1, color = "black", alpha = 0.8) +
  
  # CREATE THREE PANELS (One for each Model)
  facet_wrap(~ Model_Type, scales = "free_y") + 
  
  # Grayscale for the EJ Comparison
  scale_fill_manual(values = c(
    "Non-Disadvantaged" = "#F0F0F0", # Very Light Grey
    "Disadvantaged" = "#636363"  # Dark Grey
  )) +
  
  # Labels
  labs(
    title = "Annual Welfare Loss by Model Type and Social Vulnerabilities Status",
    subtitle = "Comparing Disadvantaged vs. Non-Disadvantaged ZIP Codes across Hawaii Parks",
    x = "Year",
    y = "Average Annual Welfare Loss ($/Park)",
    fill = "Community Status"
  ) +
  
  # Academic Theme
  theme_bw() +
  theme(
    text = element_text(family = "serif", size = 20),
    plot.title = element_text(face = "bold", size = 20, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 20),
    
    # Grid and Panel styling
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(face = "bold", size = 20),
    
    # Legend at bottom
    legend.position = "bottom",
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8)
  )
ggsave("comparision_social_observed.png", width=12, height=10, dpi=300)



```

```{r}
library(tidyverse)
library(sf)
library(tigris)
library(ggrepel)
library(ggspatial)
library(ggpattern)

options(tigris_use_cache = TRUE)
# 1. Get Hawaii Outlines
islands_outline <- states(cb = TRUE) %>%
  filter(STUSPS == "HI")

# 2. Get ZCTAs
zip_shapes <- zctas(cb = TRUE, year = 2020) %>%
  filter(str_starts(ZCTA5CE20, "967") | str_starts(ZCTA5CE20, "968"))

# 3. Zip averages — fix zip type here
zip_averages <- final_comparison_master1 %>%
  filter(parkname == "Stay_Home") %>%
  group_by(zip) %>%
  summarise(
    loss_mdcev       = mean(loss_mdcev,        na.rm = TRUE),
    is_disadvantaged = ifelse(
      mean(is_disadvantaged_50, na.rm = TRUE) > 0.5, "Yes", "No"
    ),
    .groups = "drop"
  ) %>%
  mutate(zip = as.character(zip))   # <-- KEY FIX

# Quick check before join
message(paste("Zip averages rows:", nrow(zip_averages)))
message(paste("Zip type:", class(zip_averages$zip)))
message(paste("ZCTA5CE20 type:", class(zip_shapes$ZCTA5CE20)))

# 4. Join
map_data <- zip_shapes %>%
  inner_join(zip_averages, by = c("ZCTA5CE20" = "zip"))

message(paste("Map data rows after join:", nrow(map_data)))

# 5. Bins
map_data <- map_data %>%
  mutate(
    loss_mdcev = ifelse(is.na(loss_mdcev) | loss_mdcev > 0, 0, loss_mdcev),
    loss_bin   = cut(loss_mdcev,
      breaks = c(-Inf, -46.03, -30.09, -17.94, 0),
      labels = c(">-$100",
                 "-$100 to -$50",
                 "-$50 to -$20",
                 "0 to -$20"),
      include.lowest = TRUE,
      right          = TRUE
    )
  )

# 6. Island labels
island_labels <- tibble(
  name     = c("Oʻahu", "Maui", "Hawaiʻi", "Kauaʻi", "Molokaʻi", "Lānaʻi"),
  x_anchor = c(-157.98, -156.33, -155.55, -159.52, -157.02, -156.92),
  y_anchor = c( 21.47,   20.80,   19.60,   22.06,   21.13,   20.83),
  x_label  = c(-158.60, -155.10, -157.20, -160.50, -156.60, -157.60),
  y_label  = c( 20.90,   21.50,   19.60,   22.20,   21.60,   20.15)
) %>%
  mutate(
    shrink      = 0.4,
    x_seg_start = x_label + shrink * (x_anchor - x_label),
    y_seg_start = y_label + shrink * (y_anchor - y_label)
  )

# 7. Plot
hawaii_map <- ggplot() +
  geom_sf(data = islands_outline, fill = "gray95", color = "gray60", size = 0.4) +
  
  geom_sf_pattern(
    data             = filter(map_data, is_disadvantaged == "No"),
    aes(fill = loss_bin, pattern = is_disadvantaged),
    pattern_color    = NA,
    pattern_fill     = NA,
    pattern_density  = 0.01,
    color            = "black",
    size             = 0.1
  ) +
  
  geom_sf_pattern(
    data             = filter(map_data, is_disadvantaged == "Yes"),
    aes(fill = loss_bin, pattern = is_disadvantaged),
    pattern_color    = "grey60",
    pattern_fill     = NA,
    pattern_density  = 0.4,
    pattern_spacing  = 0.03,
    pattern_angle    = 45,
    pattern_size     = 0.6,
    color            = "grey60",
    size             = 0.2
  ) +
  
  geom_segment(
    data      = island_labels,
    aes(x = x_seg_start, y = y_seg_start, xend = x_anchor, yend = y_anchor),
    color     = "gray50",
    linewidth = 0.3
  ) +
  
  geom_point(
    data  = island_labels,
    aes(x = x_anchor, y = y_anchor),
    color = "gray30",
    size  = 1.0
  ) +
  
  # White halo
  geom_text(
    data   = island_labels,
    aes(x = x_label, y = y_label, label = name),
    size   = 12, fontface = "bold", color = "white",
    family = "Times New Roman", hjust = 0.5, vjust = 0.5
  ) +
  # Actual text
  geom_text(
    data   = island_labels,
    aes(x = x_label, y = y_label, label = name),
    size   = 12, fontface = "bold", color = "gray10",
    family = "Times New Roman", hjust = 0.5, vjust = 0.5
  ) +
  
  annotation_north_arrow(
    location    = "bl",
    which_north = "true",
    height = unit(2.0, "cm"),
    width  = unit(1.5, "cm"),
    style  = north_arrow_fancy_orienteering(
      fill        = c("gray10", "white"),
      line_col    = "gray10",
      text_col    = "gray10",
      text_family = "Times New Roman",
      text_size   = 16
    ),
    pad_x = unit(1.0, "cm"),
    pad_y = unit(0.9, "cm")
  ) +
  
  annotation_scale(
    location      = "bl",
    unit_category = "imperial",
    text_family   = "Times New Roman",
    text_cex      = 1,
    style         = "bar",
    bar_cols      = c("gray10", "white"),
    line_col      = "gray10",
    text_col      = "gray10",
    pad_x = unit(1.0, "cm"),
    pad_y = unit(0.3, "cm")
  ) +
  
  scale_fill_manual(
    values = c(
      ">-$100"        = "gray10",
      "-$100 to -$50" = "gray40",
      "-$50 to -$20"  = "gray70",
      "0 to -$20"     = "gray92"
    ),
    name     = "Avg Welfare Loss",
    na.value = "transparent",
    guide    = guide_legend(
      override.aes   = list(pattern = "none"),
      reverse        = FALSE,
      keyheight      = unit(5, "mm"),
      keywidth       = unit(5, "mm"),
      label.position = "right"
    )
  ) +
  
  scale_pattern_manual(
    values = c("No" = "none", "Yes" = "crosshatch"),
    name   = "Community Type",
    labels = c("No" = "Non-Disadvantaged", "Yes" = "Disadvantaged"),
    guide  = guide_legend(
      override.aes = list(
        fill            = c("white", "white"),
        color           = c("black", "black"),
        pattern_color   = c(NA, "black"),
        pattern_fill    = c(NA, NA),
        pattern_density = c(0.01, 0.4),
        pattern_spacing = c(0.03, 0.02),
        pattern_angle   = c(45, 45)
      ),
      keyheight      = unit(7, "mm"),
      keywidth       = unit(7, "mm"),
      label.position = "right"
    )
  ) +
  
  coord_sf(xlim = c(-161.5, -154.0), ylim = c(18.7, 22.5), expand = FALSE) +
  theme_minimal() +
  theme(
    panel.grid           = element_blank(),
    axis.text            = element_blank(),
    axis.title           = element_blank(),
    axis.ticks           = element_blank(),
    legend.position      = c(0.1, 0.15),
    legend.justification = c("left", "bottom"),
    legend.box           = "vertical",
    legend.background    = element_rect(fill = "white", color = NA),
    text                 = element_text(family = "Times New Roman"),
    plot.title           = element_text(family = "Times New Roman", face = "bold"),
    legend.title         = element_text(family = "Times New Roman",
                                        face = "bold", size = 14),
    legend.text          = element_text(family = "Times New Roman", size = 14)
  )

ggsave(
  filename = "hawaii_welfare_loss_map.png",
  plot     = hawaii_map,
  width    = 12,
  height   = 10,
  dpi      = 300,
  units    = "in",
  bg       = "white"
)
```

```{r}

library(tidyverse)
library(sf)
library(tigris)
library(ggrepel)
library(ggspatial)
library(ggpattern)

# 1. Get Hawaii Outlines
islands_outline <- states(cb = TRUE) %>%
  filter(STUSPS == "HI")

# 2. Get ZCTAs
zip_shapes <- zctas(cb = TRUE, year = 2020) %>%
  filter(str_starts(ZCTA5CE20, "967") | str_starts(ZCTA5CE20, "968"))

# 3. Zip averages
zip_averages <- final_comparison_master1 %>%
  filter(parkname == "Stay_Home") %>%
  group_by(zip) %>%
  summarise(
    loss_mdcev       = mean(loss_mdcev,        na.rm = TRUE),
    is_disadvantaged = ifelse(
      mean(is_disadvantaged_50, na.rm = TRUE) > 0.5, "Yes", "No"
    ),
    .groups = "drop"
  ) %>%
  mutate(zip = as.character(zip))

# 4. Join
map_data <- zip_shapes %>%
  inner_join(zip_averages, by = c("ZCTA5CE20" = "zip"))

# 5. Bins
map_data <- map_data %>%
  mutate(
    loss_mdcev = ifelse(is.na(loss_mdcev) | loss_mdcev > 0, 0, loss_mdcev),
    loss_bin   = cut(loss_mdcev,
      breaks = c(-Inf, -46.03, -30.09, -17.94, 0),
      labels = c(">-$100",
                 "-$100 to -$50",
                 "-$50 to -$20",
                 "0 to -$20"),
      include.lowest = TRUE,
      right          = TRUE
    )
  )

# 6. Island labels
island_labels <- tibble(
  name     = c("Oʻahu", "Maui", "Hawaiʻi", "Kauaʻi", "Molokaʻi", "Lānaʻi"),
  x_anchor = c(-157.98, -156.33, -155.55, -159.52, -157.02, -156.92),
  y_anchor = c( 21.47,   20.80,   19.60,   22.06,   21.13,   20.83),
  x_label  = c(-158.60, -155.10, -157.20, -160.50, -156.60, -157.60),
  y_label  = c( 20.90,   21.50,   19.60,   22.20,   21.60,   20.15)
) %>%
  mutate(
    shrink      = 0.4,
    x_seg_start = x_label + shrink * (x_anchor - x_label),
    y_seg_start = y_label + shrink * (y_anchor - y_label)
  )

# 7. Color map — white to dark red
hawaii_map_color <- ggplot() +
  geom_sf(data = islands_outline, fill = "gray95", color = "gray60", size = 0.4) +
  
# Non-DAC layer
geom_sf_pattern(
    data             = filter(map_data, is_disadvantaged == "No"),
    aes(fill = loss_bin, pattern = is_disadvantaged),
    pattern_color    = NA,
    pattern_fill     = "transparent",   # <-- change from NA
    pattern_density  = 0.01,
    color            = "white",
    size             = 0.1
) +

# DAC layer
geom_sf_pattern(
    data             = filter(map_data, is_disadvantaged == "Yes"),
    aes(fill = loss_bin, pattern = is_disadvantaged),
    pattern_color    = "black",
    pattern_fill     = "transparent",   # <-- change from NA
    pattern_density  = 0.4,
    pattern_spacing  = 0.03,
    pattern_angle    = 45,
    pattern_size     = 0.6,
    color            = "white",
    size             = 0.2
) +
  
  geom_segment(
    data      = island_labels,
    aes(x = x_seg_start, y = y_seg_start, xend = x_anchor, yend = y_anchor),
    color     = "gray30",
    linewidth = 0.3
  ) +
  
  geom_point(
    data  = island_labels,
    aes(x = x_anchor, y = y_anchor),
    color = "gray20",
    size  = 1.0
  ) +
  
  # White halo
  geom_text(
    data   = island_labels,
    aes(x = x_label, y = y_label, label = name),
    size   = 12, fontface = "bold", color = "white",
    family = "Times New Roman", hjust = 0.5, vjust = 0.5
  ) +
  # Actual text
  geom_text(
    data   = island_labels,
    aes(x = x_label, y = y_label, label = name),
    size   = 11, fontface = "bold", color = "gray10",
    family = "Times New Roman", hjust = 0.5, vjust = 0.5
  ) +
  
  annotation_north_arrow(
    location    = "bl",
    which_north = "true",
    height = unit(2.0, "cm"),
    width  = unit(1.5, "cm"),
    style  = north_arrow_fancy_orienteering(
      fill        = c("gray10", "white"),
      line_col    = "gray10",
      text_col    = "gray10",
      text_family = "Times New Roman",
      text_size   = 16
    ),
    pad_x = unit(1.0, "cm"),
    pad_y = unit(0.9, "cm")
  ) +
  
  annotation_scale(
    location      = "bl",
    unit_category = "imperial",
    text_family   = "Times New Roman",
    text_cex      = 1,
    style         = "bar",
    bar_cols      = c("gray10", "white"),
    line_col      = "gray10",
    text_col      = "gray10",
    pad_x = unit(1.0, "cm"),
    pad_y = unit(0.3, "cm")
  ) +
  
  # ── COLOR SCALE: white (low loss) to dark red (high loss) ─────────────────
  scale_fill_manual(
    values = c(
      ">-$100"        = "#67000d",   # darkest red   — highest loss
      "-$100 to -$50" = "#cb181d",   # medium red
      "-$50 to -$20"  = "#fc8d59",   # light orange-red
      "0 to -$20"     = "#fff5f0"    # near white    — lowest loss
    ),
    name     = "Avg Welfare Loss",
    na.value = "transparent",
    guide    = guide_legend(
      override.aes   = list(pattern = "none"),
      reverse        = FALSE,
      keyheight      = unit(6, "mm"),
      keywidth       = unit(6, "mm"),
      label.position = "right"
    )
  ) +
  
  scale_pattern_manual(
    values = c("No" = "none", "Yes" = "crosshatch"),
    name   = "Community Type",
    labels = c("No" = "Non-Disadvantaged", "Yes" = "Disadvantaged"),
    guide  = guide_legend(
      override.aes = list(
        fill            = c("white", "white"),
        color           = c("black", "black"),
        pattern_color   = c(NA, "black"),
        pattern_fill    = c(NA, NA),
        pattern_density = c(0.01, 0.4),
        pattern_spacing = c(0.03, 0.02),
        pattern_angle   = c(45, 45)
      ),
      keyheight      = unit(7, "mm"),
      keywidth       = unit(7, "mm"),
      label.position = "right"
    )
  ) +
  
  coord_sf(xlim = c(-161.5, -154.0), ylim = c(18.7, 22.5), expand = FALSE) +
  theme_minimal() +
  theme(
    panel.grid           = element_blank(),
    axis.text            = element_blank(),
    axis.title           = element_blank(),
    axis.ticks           = element_blank(),
    legend.position      = c(0.1, 0.15),
    legend.justification = c("left", "bottom"),
    legend.box           = "vertical",
    legend.background    = element_rect(fill = "white", color = NA),
    text                 = element_text(family = "Times New Roman"),
    plot.title           = element_text(family = "Times New Roman", face = "bold"),
    legend.title         = element_text(family = "Times New Roman",
                                        face = "bold", size = 14),
    legend.text          = element_text(family = "Times New Roman", size = 14)
  )

ggsave(
  filename = "hawaii_welfare_loss_map_color.png",
  plot     = hawaii_map_color,
  width    = 12,
  height   = 10,
  dpi      = 300,
  units    = "in",
  bg       = "white"
)




```

# 

```{r}

library(tidyverse)
library(sf)
library(tigris)
library(ggrepel)
library(ggspatial)
library(ggpattern)

# 1. Get Hawaii Outlines
islands_outline <- states(cb = TRUE) %>%
  filter(STUSPS == "HI")

# 2. Get ZCTAs
zip_shapes <- zctas(cb = TRUE, year = 2020) %>%
  filter(str_starts(ZCTA5CE20, "967") | str_starts(ZCTA5CE20, "968"))

plot_data <- final_comparison_master1 %>%
  group_by(zip, parkname, year,is_disadvantaged_50) %>%
  filter(parkname!="Stay_Home")%>%
  summarise(
    ObservedRUM = mean(loss_rum_observed, na.rm = TRUE),
    FrequencyRUM = mean(loss_rum_52wk, na.rm = TRUE),
    .groups = "drop"
  )


# 3. Zip averages
zip_averages <- plot_data %>%
  # Step 1: Sum welfare across all sites within each zip and year
  group_by(zip, year) %>%
  summarise(
    loss_rum_observed = sum(ObservedRUM, na.rm = TRUE),
    is_disadvantaged  = ifelse(
      mean(is_disadvantaged_50, na.rm = TRUE) > 0.5, "Yes", "No"
    ),
    .groups = "drop"
  ) %>%
  
  # Step 2: Average the yearly sums across all years
  group_by(zip) %>%
  summarise(
    loss_rum_observed = mean(loss_rum_observed, na.rm = TRUE),
    is_disadvantaged  = ifelse(          # <-- uses is_disadvantaged not is_disadvantaged_50
      mean(is_disadvantaged == "Yes", na.rm = TRUE) > 0.5, "Yes", "No"
    ),
    .groups = "drop"
  ) %>%
  mutate(zip = as.character(zip))

# 4. Join
map_data <- zip_shapes %>%
  inner_join(zip_averages, by = c("ZCTA5CE20" = "zip"))

# 5. Bins
map_data <- map_data %>%
  mutate(
    loss_rum_observed = ifelse(is.na(loss_rum_observed) | loss_rum_observed > 0, 0, loss_rum_observed),
    loss_bin   = cut(loss_rum_observed,
      breaks = c(-Inf, -400, -300, -200, -100),
      labels = c(">-$400",
                 "-$400 to -$300",
                 "-$300 to -$200",
                 "-$100 to -$200"),
      include.lowest = TRUE,
      right          = TRUE
    )
  )

# 6. Island labels
island_labels <- tibble(
  name     = c("Oʻahu", "Maui", "Hawaiʻi", "Kauaʻi", "Molokaʻi", "Lānaʻi"),
  x_anchor = c(-157.98, -156.33, -155.55, -159.52, -157.02, -156.92),
  y_anchor = c( 21.47,   20.80,   19.60,   22.06,   21.13,   20.83),
  x_label  = c(-158.60, -155.10, -157.20, -160.50, -156.60, -157.60),
  y_label  = c( 20.90,   21.50,   19.60,   22.20,   21.60,   20.15)
) %>%
  mutate(
    shrink      = 0.4,
    x_seg_start = x_label + shrink * (x_anchor - x_label),
    y_seg_start = y_label + shrink * (y_anchor - y_label)
  )

# 7. Color map — white to dark red
hawaii_map_color <- ggplot() +
  geom_sf(data = islands_outline, fill = "gray95", color = "gray60", size = 0.4) +
  
# Non-DAC layer
geom_sf_pattern(
    data             = filter(map_data, is_disadvantaged == "No"),
    aes(fill = loss_bin, pattern = is_disadvantaged),
    pattern_color    = NA,
    pattern_fill     = "transparent",   # <-- change from NA
    pattern_density  = 0.01,
    color            = "white",
    size             = 0.1
) +

# DAC layer
geom_sf_pattern(
    data             = filter(map_data, is_disadvantaged == "Yes"),
    aes(fill = loss_bin, pattern = is_disadvantaged),
    pattern_color    = "black",
    pattern_fill     = "transparent",   # <-- change from NA
    pattern_density  = 0.4,
    pattern_spacing  = 0.03,
    pattern_angle    = 45,
    pattern_size     = 0.6,
    color            = "white",
    size             = 0.2
) +
  
  geom_segment(
    data      = island_labels,
    aes(x = x_seg_start, y = y_seg_start, xend = x_anchor, yend = y_anchor),
    color     = "gray30",
    linewidth = 0.3
  ) +
  
  geom_point(
    data  = island_labels,
    aes(x = x_anchor, y = y_anchor),
    color = "gray20",
    size  = 1.0
  ) +
  
  # White halo
  geom_text(
    data   = island_labels,
    aes(x = x_label, y = y_label, label = name),
    size   = 12, fontface = "bold", color = "white",
    family = "Times New Roman", hjust = 0.5, vjust = 0.5
  ) +
  # Actual text
  geom_text(
    data   = island_labels,
    aes(x = x_label, y = y_label, label = name),
    size   = 11, fontface = "bold", color = "gray10",
    family = "Times New Roman", hjust = 0.5, vjust = 0.5
  ) +
  
  annotation_north_arrow(
    location    = "bl",
    which_north = "true",
    height = unit(2.0, "cm"),
    width  = unit(1.5, "cm"),
    style  = north_arrow_fancy_orienteering(
      fill        = c("gray10", "white"),
      line_col    = "gray10",
      text_col    = "gray10",
      text_family = "Times New Roman",
      text_size   = 16
    ),
    pad_x = unit(1.0, "cm"),
    pad_y = unit(0.9, "cm")
  ) +
  
  annotation_scale(
    location      = "bl",
    unit_category = "imperial",
    text_family   = "Times New Roman",
    text_cex      = 1,
    style         = "bar",
    bar_cols      = c("gray10", "white"),
    line_col      = "gray10",
    text_col      = "gray10",
    pad_x = unit(1.0, "cm"),
    pad_y = unit(0.3, "cm")
  ) +
  
  # ── COLOR SCALE: white (low loss) to dark red (high loss) ─────────────────
  scale_fill_manual(
    values = c(
      ">-$400"        = "#67000d",   # darkest red   — highest loss
      "-$400 to -$300" = "#cb181d",   # medium red
      "-$300 to -$200"  = "#fc8d59",   # light orange-red
      "-$100 to -$200"     = "#fff5f0"    # near white    — lowest loss
    ),
    name     = "Avg Welfare Loss",
    na.value = "transparent",
    guide    = guide_legend(
      override.aes   = list(pattern = "none"),
      reverse        = FALSE,
      keyheight      = unit(6, "mm"),
      keywidth       = unit(6, "mm"),
      label.position = "right"
    )
  ) +
  
  scale_pattern_manual(
    values = c("No" = "none", "Yes" = "crosshatch"),
    name   = "Community Type",
    labels = c("No" = "Non-Disadvantaged", "Yes" = "Disadvantaged"),
    guide  = guide_legend(
      override.aes = list(
        fill            = c("white", "white"),
        color           = c("black", "black"),
        pattern_color   = c(NA, "black"),
        pattern_fill    = c(NA, NA),
        pattern_density = c(0.01, 0.4),
        pattern_spacing = c(0.03, 0.02),
        pattern_angle   = c(45, 45)
      ),
      keyheight      = unit(7, "mm"),
      keywidth       = unit(7, "mm"),
      label.position = "right"
    )
  ) +
  
  coord_sf(xlim = c(-161.5, -154.0), ylim = c(18.7, 22.5), expand = FALSE) +
  theme_minimal() +
  theme(
    panel.grid           = element_blank(),
    axis.text            = element_blank(),
    axis.title           = element_blank(),
    axis.ticks           = element_blank(),
    legend.position      = c(0.1, 0.15),
    legend.justification = c("left", "bottom"),
    legend.box           = "vertical",
    legend.background    = element_rect(fill = "white", color = NA),
    text                 = element_text(family = "Times New Roman"),
    plot.title           = element_text(family = "Times New Roman", face = "bold"),
    legend.title         = element_text(family = "Times New Roman",
                                        face = "bold", size = 14),
    legend.text          = element_text(family = "Times New Roman", size = 14)
  )

ggsave(
  filename = "hawaii_welfare_loss_map_colorRUM.png",
  plot     = hawaii_map_color,
  width    = 12,
  height   = 10,
  dpi      = 300,
  units    = "in",
  bg       = "white"
)



```

# Map

```{r}

print(hawaii_map_color)


print(hawaii_map)
```

# Sensitivity Test

```{r}
get_price_coef <- function(mle, yr, model_name) {
  
  tryCatch({
    # mlogit coefficients
    beta  <- coef(mle)["price"]
    
    # variance-covariance matrix
    se    <- sqrt(vcov(mle)["price", "price"])
    tibble(
      year       = yr,
      model      = model_name,
      beta_price = beta,
      se_price   = se,
      ci_low     = beta - 1.96 * se,
      ci_high    = beta + 1.96 * se
    )
  }, error = function(e) {
    message(paste("Failed for", model_name, yr, ":", e$message))
    tibble(
      year       = yr,
      model      = model_name,
      beta_price = NA_real_,
      se_price   = NA_real_,
      ci_low     = NA_real_,
      ci_high    = NA_real_
    )
  })
}



# PRice Stability
price_stability_all <- map_df(2018:2023, function(yr) {
  
  obs_key  <- paste0("Year_", yr)   
  freq_key <- paste0("52wk_", yr)   
  
  obs_mle  <- model_list_obs[[obs_key]]
  freq_mle <- model_list_52[[freq_key]]
  
  bind_rows(
    if (!is.null(obs_mle))
      get_price_coef(obs_mle,  yr, "Observed RUM"),
    if (!is.null(freq_mle))
      get_price_coef(freq_mle, yr, "Frequency RUM")
  )
})

print(price_stability_all)
year_pairs <- combn(2018:2023, 2, simplify = FALSE)
# Pairwise welfare stability — all three models 
test_year_pairs_model <- function(yr1, yr2, loss_col) {
  
  d1 <- final_comparison_master %>%
    filter(year == yr1, parkname != "Stay_Home") %>%
    pull({{ loss_col }})
  
  d2 <- final_comparison_master %>%
    filter(year == yr2, parkname != "Stay_Home") %>%
    pull({{ loss_col }})
  
  t_result <- t.test(d1, d2)
  
  tibble(
    year1   = yr1,
    year2   = yr2,
    mean1   = round(mean(d1, na.rm = TRUE), 2),
    mean2   = round(mean(d2, na.rm = TRUE), 2),
    t_stat  = round(t_result$statistic, 3),
    p_value = round(t_result$p.value,   4),
    stable  = t_result$p.value > 0.05
  )
}

stability_all <- bind_rows(
  map_df(year_pairs, ~test_year_pairs_model(.x[1], .x[2], loss_mdcev)) %>%
    mutate(model = "MDCEV"),
  map_df(year_pairs, ~test_year_pairs_model(.x[1], .x[2], loss_rum_observed)) %>%
    mutate(model = "Observed RUM"),
  map_df(year_pairs, ~test_year_pairs_model(.x[1], .x[2], loss_rum_52wk)) %>%
    mutate(model = "Frequency RUM")
)




# Convergent validity across models — do they agree on rankings? 
site_welfare <- final_comparison_master %>%
    filter(parkname != "Stay_Home") %>%
    group_by(parkname) %>%
    summarise(
        mean_mdcev = mean(loss_mdcev,        na.rm = TRUE),
        mean_obs   = mean(loss_rum_observed, na.rm = TRUE),
        mean_52wk  = mean(loss_rum_52wk,     na.rm = TRUE),
        .groups    = "drop"
    )

# Spearman rank correlations across models
cross_model_cor <- site_welfare %>%
    select(mean_mdcev, mean_obs, mean_52wk) %>%
    cor(method = "spearman", use = "complete.obs")

colnames(cross_model_cor) <- c("MDCEV", "Observed RUM", "Frequency RUM")
rownames(cross_model_cor) <- c("MDCEV", "Observed RUM", "Frequency RUM")

print(round(cross_model_cor, 3))

# Transfer error across models 
# Ji et al. transfer error as:
# TE = |predicted - actual| / |actual| × 100

cross_model_te <- site_welfare %>%
    mutate(
        # Use MDCEV to predict each RUM
        te_mdcev_to_obs  = abs((mean_mdcev - mean_obs)   / mean_obs)  * 100,
        te_mdcev_to_52wk = abs((mean_mdcev - mean_52wk)  / mean_52wk) * 100,
        # Use observed RUM to predict frequency RUM
        te_obs_to_52wk   = abs((mean_obs   - mean_52wk)  / mean_52wk) * 100
    ) %>%
    summarise(
        # Mean transfer error across sites
        mean_te_mdcev_obs  = round(mean(te_mdcev_to_obs,  na.rm = TRUE), 1),
        mean_te_mdcev_52wk = round(mean(te_mdcev_to_52wk, na.rm = TRUE), 1),
        mean_te_obs_52wk   = round(mean(te_obs_to_52wk,   na.rm = TRUE), 1),
        # Median transfer error
        med_te_mdcev_obs   = round(median(te_mdcev_to_obs,  na.rm = TRUE), 1),
        med_te_mdcev_52wk  = round(median(te_mdcev_to_52wk, na.rm = TRUE), 1),
        med_te_obs_52wk    = round(median(te_obs_to_52wk,   na.rm = TRUE), 1)
    )

print(cross_model_te)

# Equality tests — are model welfare estimates significantly different from each other? 
# (Ji et al. Section 4)

# Paired t-test: MDCEV vs Observed RUM
t_mdcev_obs <- t.test(
    final_comparison_master$loss_mdcev,
    final_comparison_master$loss_rum_observed,
    paired = TRUE
)

# Paired t-test: MDCEV vs Frequency RUM
t_mdcev_52wk <- t.test(
    final_comparison_master$loss_mdcev,
    final_comparison_master$loss_rum_52wk,
    paired = TRUE
)

# Paired t-test: Observed RUM vs Frequency RUM
t_obs_52wk <- t.test(
    final_comparison_master$loss_rum_observed,
    final_comparison_master$loss_rum_52wk,
    paired = TRUE
)

equality_tests <- tibble(
    comparison     = c("MDCEV vs Observed RUM",
                       "MDCEV vs Frequency RUM",
                       "Observed RUM vs Frequency RUM"),
    mean_diff      = c(round(t_mdcev_obs$estimate,  2),
                       round(t_mdcev_52wk$estimate, 2),
                       round(t_obs_52wk$estimate,   2)),
    t_stat         = c(round(t_mdcev_obs$statistic,  3),
                       round(t_mdcev_52wk$statistic, 3),
                       round(t_obs_52wk$statistic,   3)),
    p_value        = c(round(t_mdcev_obs$p.value,  4),
                       round(t_mdcev_52wk$p.value, 4),
                       round(t_obs_52wk$p.value,   4)),
    significantly_different = c(
        t_mdcev_obs$p.value  < 0.05,
        t_mdcev_52wk$p.value < 0.05,
        t_obs_52wk$p.value   < 0.05
    )
)

print(equality_tests)

# Transfer error by year pair AND model pair 
# Ji et al.'s most comprehensive test — applies both dimensions

te_full <- final_comparison_master %>%
    filter(parkname != "Stay_Home", year!=2020, year!=2021) %>%
    group_by(parkname, year) %>%
    summarise(
        mean_mdcev = mean(loss_mdcev,        na.rm = TRUE),
        mean_obs   = mean(loss_rum_observed, na.rm = TRUE),
        mean_52wk  = mean(loss_rum_52wk,     na.rm = TRUE),
        .groups    = "drop"
    ) %>%
    full_join(
        final_comparison_master %>%
        
            filter(parkname != "Stay_Home") %>%
            group_by(parkname, year) %>%
            summarise(
                pred_mdcev = mean(loss_mdcev,        na.rm = TRUE),
                pred_obs   = mean(loss_rum_observed, na.rm = TRUE),
                pred_52wk  = mean(loss_rum_52wk,     na.rm = TRUE),
                .groups    = "drop"
            ) %>%
            rename(year2 = year),
        by = "parkname"
    ) %>%
    filter(year != year2) %>%
    mutate(
        # Within model temporal transfer error
        te_mdcev_time = abs((pred_mdcev - mean_mdcev) / mean_mdcev) * 100,
        te_obs_time   = abs((pred_obs   - mean_obs)   / mean_obs)   * 100,
        te_52wk_time  = abs((pred_52wk  - mean_52wk)  / mean_52wk)  * 100,
        # Cross model transfer error (use MDCEV from year1 to predict RUM in year2)
        te_mdcev_to_obs_cross  = abs((pred_mdcev - mean_obs)   / mean_obs)   * 100,
        te_mdcev_to_52wk_cross = abs((pred_mdcev - mean_52wk)  / mean_52wk)  * 100
    ) %>%
    group_by(year, year2) %>%
    summarise(
        te_mdcev       = round(mean(te_mdcev_time,na.rm = TRUE), 1),
        te_obs         = round(mean(te_obs_time,na.rm = TRUE), 1),
        te_52wk        = round(mean(te_52wk_time,na.rm = TRUE), 1),
        te_cross_mdcev_obs  = round(mean(te_mdcev_to_obs_cross,  na.rm = TRUE), 1),
        te_cross_mdcev_52wk = round(mean(te_mdcev_to_52wk_cross, na.rm = TRUE), 1),
        .groups        = "drop"
    )

print(te_full, n = 30)

# Summary table matching Ji et al. Table format
summary_ji <- tibble(
    Test               = c(
        "Mean Spearman Rank Correlation",
        "Mean Transfer Error — MDCEV",
        "Mean Transfer Error — Observed RUM",
        "Mean Transfer Error — Frequency RUM",
        "% Year Pairs Stable — MDCEV",
        "% Year Pairs Stable — Observed RUM",
        "% Year Pairs Stable — Frequency RUM"
    ),
    Value = c(
        round(mean(c(cross_model_cor[1,2],
                     cross_model_cor[1,3],
                     cross_model_cor[2,3])), 3),
        round(mean(te_full$te_mdcev,  na.rm = TRUE), 1),
        round(mean(te_full$te_obs,    na.rm = TRUE), 1),
        round(mean(te_full$te_52wk,   na.rm = TRUE), 1),
        round(mean(stability_all$stable[stability_all$model == "MDCEV"])         * 100, 1),
        round(mean(stability_all$stable[stability_all$model == "Observed RUM"])  * 100, 1),
        round(mean(stability_all$stable[stability_all$model == "Frequency RUM"]) * 100, 1)
    )
)

print(summary_ji)


```
