---
title: "rum"

format:
  html:
    code-fold: false        # Enables dropdown for code
    code-tools: true       # (Optional) Adds buttons like "Show Code"
    code-summary: "Show code"  # (Optional) Custom label for dropdown
    toc: true
    toc-location: left
    page-layout: full
editor: visual
---

# Rum

This script will run a RUM model on our data assuming a stay at home option and compare it to our result of the MDCEV

## Function

This function calculate the rum welfare manual and is used for each year to get a comparison to the MDCEV.

```{r}

calc_rum_welfare_manual <- function(target_park) {
  
  beta_price <- coef(rum_mle)["price"]
  
  # 1. Baseline Logsum
  ls_base <- df_math %>%
    group_by(trip_id) %>%
    summarise(ls_base = log(sum(exp(V))), .groups = "drop")
  
  # 2. Policy Logsum (Removing the park one by one and then use the consumer surplus using closed form based on the marginal utility of income
  ls_policy <- df_math %>%
    filter(park_name != target_park) %>%
    group_by(trip_id) %>%
    summarise(ls_policy = log(sum(exp(V))), .groups = "drop")
  
  # 3. Calculate CV for EVERY individual trip
  comparison <- left_join(ls_base, ls_policy, by = "trip_id") %>%
    mutate(cv = (ls_policy - ls_base) / -beta_price) # this is trying to compare to the MDCEV where we close and re-estimate
  
  # 4. Calculate Statistics across all trips
  # We use the standard error of the mean for the CIs
  res <- comparison %>%
    summarise(
      mean_cv_rum = mean(cv, na.rm = TRUE),
      std_dev_rum = sd(cv, na.rm = TRUE),
      n = n(),
      # Standard Error = SD / sqrt(n)
      se_rum = std_dev_rum / sqrt(n),
      # 95% CI calculation (Mean +/- 1.96 * SE) I do this to compare the the same CI in the MDCEV
      ci_lo_2.5_rum = mean_cv_rum - (1.96 * se_rum),
      ci_hi_97.5_rum = mean_cv_rum + (1.96 * se_rum)
    )
  
  return(res)
}

```

### 2018

Need to run the same data for the MDCEV which gives us final_data_2018. So going to clip the code from the file R script high cost.

The first code will produce final_data_yyyy by year and then should be able to run all the rest of the data

```{r}


library(dplyr)
library(tidyr)
 library(haven)
library(readr)

# Load datasets
reservation_data <- read_dta("reservation_data_merged_id_complete.dta")
intra_island_distance <- read.csv("intra_island_distance2.csv")

# Add year
reservation_data$year <- as.numeric(format(as.Date(reservation_data$CheckinDate, "%Y-%m-%d"), "%Y"))

reservation_data <- reservation_data %>%
  mutate(zip = ifelse(zip == 96812, 96813, zip))

# Group and compute yearly averages
avg_days_per_year <- reservation_data %>%
  group_by(year, id) %>%
  summarize(avg_days_visited = mean(days_visited, na.rm = TRUE)) %>%
  ungroup()

avg_days_2018 <- avg_days_per_year %>% filter(year == 2018)
avg_days_2019 <- avg_days_per_year %>% filter(year == 2019)
avg_days_2020 <- avg_days_per_year %>% filter(year == 2020)
avg_days_2021 <- avg_days_per_year %>% filter(year == 2021)
avg_days_2022 <- avg_days_per_year %>% filter(year == 2022)
avg_days_2023 <- avg_days_per_year %>% filter(year == 2023)


# Function to process data for a single year
process_year <- function(year_input) {
  data_year <- reservation_data %>% filter(year == year_input)
  
  unique_visits <- data_year %>%
    filter(!is.na(CheckinDate)) %>%
    distinct(id, parkname, CheckinDate) %>%
    group_by(id, parkname) %>%
    summarize(VisitCount = n(), .groups = "drop")
  
  all_combinations <- expand.grid(
    id = unique(data_year$id),
    parkname = unique(data_year$parkname),
    stringsAsFactors = FALSE
  )
  
  visit_data <- all_combinations %>%
    left_join(unique_visits, by = c("id", "parkname")) %>%
    mutate(VisitCount = ifelse(is.na(VisitCount), 0, VisitCount))
  
  person_info <- data_year %>%
    select(id, Permittee, zip, Zip, NoOfGuests, days_visited, year) %>%
    distinct()
  
  long_resdata_count <- left_join(visit_data, person_info, by = "id")
  
  long_resdata_count <- long_resdata_count %>%
    mutate(island_park = case_when(
      parkname == "Ainapo Cabin" ~ "Hawaii",
      parkname == "Hipalau" ~ "Kauai",
      parkname == "Kaluahaulu" ~ "Kauai",
      parkname == "Kamananui Trail" ~ "Oahu",
      parkname == "Kaunala Trail" ~ "Oahu",
      parkname == "Kawaikoi" ~ "Kauai",
      parkname == "Keanakolu Ranger & Bunkhouse Cabins" ~ "Hawaii",
      parkname == "Kuaokala Trail" ~ "Oahu",
      parkname == "Kulanaahane Trail" ~ "Oahu",
      parkname == "Kuliouou Trail" ~ "Oahu",
      parkname == "Lonomea" ~ "Kauai",
      parkname == "Maakua Ridge Trail" ~ "Oahu",
      parkname == "Manana Trail" ~ "Oahu",
      parkname == "Peacock Flats Campsite" ~ "Oahu",
      parkname == "Sugi Grove" ~ "Kauai",
      parkname == "Waialae Cabin Campsite" ~ "Kauai",
      parkname == "Waikoali" ~ "Kauai",
      parkname == "Waikolu" ~ "Molokai",
      parkname == "Waimano Trail" ~ "Oahu",
      parkname == "Waimanu Campsite" ~ "Hawaii",
      parkname == "Wiliwili" ~ "Kauai",
      parkname == "Wiliwilinui Ridge Trail" ~ "Oahu",
      TRUE ~ NA_character_
    ))
  
  long_resdata_count <- long_resdata_count[order(long_resdata_count$id, long_resdata_count$parkname), ]
  long_resdata_count$parkname_numeric <- as.integer(factor(long_resdata_count$parkname, levels = unique(long_resdata_count$parkname)))
  
  long_resdata_count <- long_resdata_count %>%
    distinct(id, parkname, .keep_all = TRUE)
  
  final_data <- long_resdata_count %>%
    left_join(
      intra_island_distance %>%
        select(zip, parkname, income, residents_distance_meters, residents_duration, island_live),
      by = c("zip", "parkname")
    )
  
  final_data <- final_data %>%
    group_by(id) %>%
    mutate(visited_another_island = ifelse(any(island_park != island_live & VisitCount > 0), 1, 0)) %>%
    ungroup()
  
  final_data$travel_distance_km <- (final_data$residents_distance_meters / 1000) * 2
  
  
  return(final_data)
}

# Run the function for each year
final_data_2018 <- process_year(2018)
final_data_2019 <- process_year(2019)
final_data_2020 <- process_year(2020)
final_data_2021 <- process_year(2021)
final_data_2022 <- process_year(2022)
final_data_2023 <- process_year(2023)

# Add correct avg_days_visited for each year by joining on id

final_data_2018 <- final_data_2018 %>%
  left_join(avg_days_2018 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2019 <- final_data_2019 %>%
  left_join(avg_days_2019 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2020 <- final_data_2020 %>%
  left_join(avg_days_2020 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2021 <- final_data_2021 %>%
  left_join(avg_days_2021 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2022 <- final_data_2022 %>%
  left_join(avg_days_2022 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

final_data_2023 <- final_data_2023 %>%
  left_join(avg_days_2023 %>% select(id, avg_days_visited), by = "id") %>%
  mutate(days_visited = avg_days_visited)

## Travel cost per year calculations ##


##Adjusted for inflation
gas_prices <- data.frame(
  year = c(2018, 2019, 2020, 2021, 2022, 2023),
  gas_price = c(4.68, 4.49, 4.09, 4.75, 5.81, 5.09)
)

mean_gas_price <- (4.68+4.49+4.09+4.75+5.81+5.09)/6


# Inputs based on AAA Brochure 2018
fuel_efficiency_mpg <- 25
gas_price_hawaii_2018 <- 4.68  # in USD/gallon
maintenance_per_mile <- 0.0858*1.2816  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6202*1.2816  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2018 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://exchange.aaa.com/wp-content/uploads/2018/09/18-0090_2018-Your-Driving-Costs-Brochure_FNL-Lo-5-2.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2018$price <- (final_data_2018$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2018$income * 0.75) / 2080 * final_data_2018$travel_distance_km / 50)


# Inputs based on AAA Brochure 2019
fuel_efficiency_mpg <- 25
gas_price_hawaii_2019 <- 4.49  # in USD/gallon
maintenance_per_mile <- 0.0960*1.262  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6845*1.262  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2019 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://exchange.aaa.com/wp-content/uploads/2019/09/AAA-Your-Driving-Costs-2019.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2019$price <- (final_data_2019$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2019$income * 0.75) / 2080 * final_data_2019$travel_distance_km / 50)

# Inputs based on AAA Brochure 2020
fuel_efficiency_mpg <- 25
gas_price_hawaii_2020 <- 4.09  # in USD/gallon
maintenance_per_mile <- 0.1011*1.2314  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6831*1.2314  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2020 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://newsroom.aaa.com/wp-content/uploads/2020/12/2020-Your-Driving-Costs-Brochure-Interactive-FINAL-12-9-20.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2020$price <- (final_data_2020$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2020$income * 0.75) / 2080 * final_data_2020$travel_distance_km / 50)

# Inputs based on AAA Brochure 2021
fuel_efficiency_mpg <- 25
gas_price_hawaii_2021 <- 4.75  # in USD/gallon
maintenance_per_mile <- 0.1043*1.12144  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6633*1.12144  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2021 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

#https://newsroom.aaa.com/wp-content/uploads/2021/08/2021-YDC-Brochure-Live.pdf

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2021$price <- (final_data_2021$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2021$income * 0.75) / 2080 * final_data_2021$travel_distance_km / 50)

# Inputs based on AAA Brochure 2022
fuel_efficiency_mpg <- 25
gas_price_hawaii_2022 <- 5.81  # in USD/gallon
maintenance_per_mile <- 0.1064*1.113  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 6599*1.113  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2022 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2022$price <- (final_data_2022$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2022$income * 0.75) / 2080 * final_data_2022$travel_distance_km / 50)

#https://newsroom.aaa.com/wp-content/uploads/2022/08/2022-YDC-Costs-Break-Out-by-Category.pdf

# Inputs based on AAA Brochure 2023
fuel_efficiency_mpg <- 25
gas_price_hawaii_2023 <- 5.09  # in USD/gallon
maintenance_per_mile <- 0.1085*1.0618  # in USD/mile 2025 price
annual_miles <- 15000
annual_ownership_cost <- 7542*1.0618  # USD 2025 price

# Calculate total cost per mile
cost_per_mile <- (
  (1 / fuel_efficiency_mpg * gas_price_hawaii_2023 * annual_miles) +  # fuel cost
    (maintenance_per_mile * annual_miles) +                             # maintenance
    annual_ownership_cost                                               # ownership
) / annual_miles  # divide by miles to get per-mile cost

# Show result
cost_per_mile

miles_to_km <- 1.60934

cost_per_km <- cost_per_mile / miles_to_km
cost_per_km

final_data_2023$price <- (final_data_2023$travel_distance_km * cost_per_km) / 3 +
  ((final_data_2023$income * 0.75) / 2080 * final_data_2023$travel_distance_km / 50)

#https://newsroom.aaa.com/wp-content/uploads/2023/08/YDC-Brochure_2023-FINAL-8.30.23-.pdf


final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))


final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(parkname == "Ainapo Cabin", (30*days_visited)/3, 0),
         price = price + ifelse(parkname == "Hipalau", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaluahaulu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kamananui Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kaunala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kawaikoi", (20*days_visited)/3, 0),
         price = price + ifelse(parkname == "Keanakolu Ranger & Bunkhouse Cabins", (60*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuaokala Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kulanaahane Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Kuliouou Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Lonomea", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Maakua Ridge Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Manana Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Peacock Flats Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Sugi Grove", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waialae Cabin Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikoali", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waikolu", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimano Trail", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Waimanu Campsite", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwili", (12*days_visited)/3, 0),
         price = price + ifelse(parkname == "Wiliwilinui Ridge Trail", (12*days_visited)/3, 0))

#########Value of travel time in air#########

### BIG ISLAND ###
##Insert value of spare time of travelling with flight to BI if live in Oahu (90 minutes before flight + 50 minutes flight time)
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu",
                                (((income * 0.75) / 2080) * 2.33) * 2, 0))

##Insert value of spare time of travelling with flight to BI if live in Kauai (90 minutes before flight + 55 minutes flight time)
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai",
                                (((income * 0.75) / 2080) * 2.42) * 2, 0))

##Insert value of spare time of travelling with flight to BI if live in Molokai (90 minutes before flight + 1.5 hour flight time one way)
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii",
                                (((income * 0.75) / 2080) * 2.33) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii",
                                (((income * 0.75) / 2080) * 2.42) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai",
                                (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2.33) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Maui",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 2) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 2.17) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 4.34) * 2, 0))

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai",
                                (((income * 0.75) / 2080) * 4.34) * 2, 0))


#########Value of travel time in air#########

### BIG ISLAND ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

#########Value of travel time in air#########

### BIG ISLAND ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

##

### BIG ISLAND ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

##

### BIG ISLAND ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))

##

### BIG ISLAND ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### OAHU ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### KAUAI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 2.42) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MOLOKAI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (((income * 0.75) / 2080) * 3.5) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (((income * 0.75) / 2080) * 3.5) * 2, 0))

### MAUI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", (((income * 0.75) / 2080) * 2.33) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", (((income * 0.75) / 2080) * 2.17) * 2, 0))

### LANAI ###
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", (((income * 0.75) / 2080) * 2) * 2, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", (((income * 0.75) / 2080) * 2.17) * 2, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", (((income * 0.75) / 2080) * 4.34) * 2, 0))



## Flight costs 2018
#########Flight cost - 2018 #########

## BIG ISLAND ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 354.0447586, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 368.0753954, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 308.7662656, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 273.2598837, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 354.0447586 + 137.8000852, 0))  # via Oahu

## OAHU ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 354.0447586, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 269.4129138, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 112.5000815, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 291.0965957, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 137.8000852, 0))

## KAUAI ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 368.0753954, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 269.4129138, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 317.7951903, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 359.4733617, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 336.1283623, 0))

## MOLOKAI ##
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 308.7662656, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 112.5000815, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 317.7951903, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 118.8094464, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 137.8000852 + 112.5000815, 0))  # via Oahu

## Flight costs 2019
## BIG ISLAND ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 348.630216, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 362.446277, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 304.044185, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 269.080816, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 348.630216 + 135.692656, 0))  # via Oahu

## OAHU ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 348.630216, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 265.292679, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 110.779575, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 286.644744, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 135.692656, 0))

## KAUAI ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 362.446277, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 265.292679, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 312.935027, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 353.975798, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 330.987822, 0))

## MOLOKAI ##
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 304.044185, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 110.779575, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 312.935027, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 118.809446, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 135.692656 + 110.779575, 0))  # via Oahu


## Flight costs 2020
## BIG ISLAND ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 340.176901, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 353.657960, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.671957, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 262.556352, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 340.176901 + 132.402485, 0))  # via Oahu

## OAHU ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 340.176901, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 258.860067, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 108.093477, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 279.694404, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.402485, 0))

## KAUAI ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 353.657960, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 258.860067, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 305.347220, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 345.392866, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 322.962286, 0))

## MOLOKAI ##
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.671957, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 108.093477, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 305.347220, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 108.093477 + 279.6944039, 0), # via Oahu
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 132.402485 + 108.093477, 0))  # via Oahu

##Flight costs 2021
## Flight costs 2021
## BIG ISLAND ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 309.800214, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 322.077459, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 239.110927, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 309.800214 + 132.4024851, 0))  # via Oahu #Use value for 2020 for Oahu to Lanai

## OAHU ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 309.800214, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 235.744708, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 126.263151, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 254.718607, 0), 
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.4024851, 0)) #Use value for 2020

## KAUAI ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 322.077459, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 235.744708, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 389.441668, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 314.550411, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 235.744708 + 132.4024851, 0))  # via Oahu #Use value for 2020 for Oahu to Lanai

## MOLOKAI ##
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 120.699975, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 389.441668, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 120.699975 + 254.718607, 0), # via Oahu
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 126.263151 + 120.699975, 0))  # via Oahu

##Flight cost 2022
## Flight costs 2022

## BIG ISLAND ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 307.468646, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 319.6534918, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 237.311369, 0),
         price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 307.468646 + 132.4024851, 0))  # via Oahu

## OAHU ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 307.468646, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 233.970485, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 169.716317, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 252.801585, 0),
         price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.4024851, 0))  #Use value for 2020

## KAUAI ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 319.6534918, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 233.970485, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 589.133955, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 312.1830927, 0),
         price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 233.970485 + 132.4024851, 0))  # via Oahu #Use value for 2020

## MOLOKAI ##
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.6719565, 0), #Use value for 2020
         price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 169.716317, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 275.9878642, 0),
         price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 169.716317 + 252.8015848, 0),  # via Oahu
         price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 169.716317 + 132.4024851, 0))  # via Oahu

##Flight cost 2023
## HAWAII ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", 293.324536, 0)) %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", 304.948857, 0)) %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", 296.6719565, 0)) %>% #Use value for 2020
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Maui", 226.39462, 0)) %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Lanai", 132.4024851 + 293.324536, 0))

## OAHU ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", 293.324536, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", 223.207422, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", 169.464607, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Maui", 241.172258, 0)) %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Lanai", 132.4024851, 0)) #Use value for 2020

## KAUAI ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", 304.948857, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", 223.207422, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", 500.736032, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Maui", 297.822109, 0)) %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Lanai", 223.207422 + 132.4024851, 0))

## MOLOKAI ##
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", 296.6719565, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", 169.464607, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", 263.291927, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Maui", 241.172258 + 169.464607, 0)) %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Lanai", 132.4024851 + 169.464607, 0))



#########Car rental cost#########

### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2018 <- final_data_2018 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2019 <- final_data_2019 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2020 <- final_data_2020 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2021 <- final_data_2021 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2022 <- final_data_2022 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


### BIG ISLAND ###

#On average a rental car in The Big Island costs $430 per week ($61 per day).

##Insert car rental cost in BI
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (61 * days_visited)/3.78, 0))

##Insert car rental cost in BI
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Molokai", (61 * days_visited)/3.78, 0))

### OAHU ###

#On average a rental car in Oahu is $48/day

##Insert car rental cost in Oahu
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (48 * days_visited)/3.78, 0))

##Insert car rental cost in Oahu
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Molokai", (48 * days_visited)/3.78, 0))

### Kauai ###

#On average a rental car in Kauai is $62/day

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (63 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Molokai", (63 * days_visited)/3.78, 0))

### Molokai ###

#On average a rental car in Molokai is $107/day

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (107 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (107 * days_visited)/3.78, 0))

#########Parking fees#########

### Oahu ###

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Oahu", (22 * days_visited)/3.78, 0))

### Kauai ###

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Hawaii" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Kauai", (24 * days_visited)/3.78, 0))

### Big Island ###

final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Oahu" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Kauai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))

##Insert car rental cost in Kauai
final_data_2023 <- final_data_2023 %>%
  mutate(price = price + ifelse(island_park == "Molokai" & island_live == "Hawaii", (24 * days_visited)/3.78, 0))


##Mean travel cost
summary(final_data_2018$price)

final_data_2018 <- final_data_2018 %>%
  arrange(Permittee, parkname)

final_data_2018$parkname_numeric <- as.integer(factor(final_data_2018$parkname, levels = unique(final_data_2018$parkname)))

##Mean travel cost
summary(final_data_2019$price)

final_data_2019 <- final_data_2019 %>%
  arrange(Permittee, parkname)

final_data_2019$parkname_numeric <- as.integer(factor(final_data_2019$parkname, levels = unique(final_data_2019$parkname)))

##Mean travel cost
summary(final_data_2020$price)

final_data_2020 <- final_data_2020 %>%
  arrange(Permittee, parkname)

final_data_2020$parkname_numeric <- as.integer(factor(final_data_2020$parkname, levels = unique(final_data_2020$parkname)))

##Mean travel cost
summary(final_data_2021$price)

final_data_2021 <- final_data_2021 %>%
  arrange(Permittee, parkname)

final_data_2021$parkname_numeric <- as.integer(factor(final_data_2021$parkname, levels = unique(final_data_2021$parkname)))

##Mean travel cost
summary(final_data_2022$price)

final_data_2022 <- final_data_2022 %>%
  arrange(Permittee, parkname)

final_data_2022$parkname_numeric <- as.integer(factor(final_data_2022$parkname, levels = unique(final_data_2022$parkname)))

##Mean travel cost
summary(final_data_2023$price)

final_data_2023 <- final_data_2023 %>%
  arrange(Permittee, parkname)

final_data_2023$parkname_numeric <- as.integer(factor(final_data_2023$parkname, levels = unique(final_data_2023$parkname)))

```

#### Rum

```{r setup}
library(mlogit)


# Create unique choice occasions for every visit
data_rum <- final_data_2018 %>%
  filter(VisitCount > 0) %>%
  uncount(VisitCount) %>%
  group_by(id) %>%
  mutate(choice_id = paste0(id, "_", row_number())) %>%
  ungroup() %>%
  select(choice_id, id, chosen_park = parkname) %>%
  expand_grid(parkname = unique(final_data_2018$parkname)) %>%
  mutate(choice = (chosen_park == parkname)) %>%
  left_join(final_data_2018 %>% select(id, parkname, price, travel_distance_km,income,zip), 
            by = c("id", "parkname"))

# Create Stay at Home

# Create the Stay-at-Home option for every choice occasion
outside_home <- data_rum %>%
  distinct(choice_id, id,income,zip) %>%
  mutate(
    parkname = "Stay_Home",
    choice = FALSE,      
    price = 0,
    travel_distance_km = 0
  )

# Combine with your existing data
data_rum<- bind_rows(data_rum, outside_home)

# Re-format for mlogit
df_rum <- dfidx(data_rum, 
                        idx = list("choice_id", "parkname"), 
                        shape = "long")

# Simple Multinomial Logit
rum_mle <- mlogit(choice ~ price  | 1, data = df_rum)
summary(rum_mle)



# Calculate the 'Logsum' (Total Utility) for the status quo
logsum_base <- logsum(rum_mle)

# 1. Convert df_rum to a standard data frame for our math
df_math <- as.data.frame(df_rum)

# 2. Extract the IDs and Park names from the 'idx' column
# idx(df_rum, 1) gets the 'choice_id'
# idx(df_rum, 2) gets the 'parkname'
df_math$trip_id   <- as.character(idx(df_rum, 1))
df_math$park_name <- as.character(idx(df_rum, 2))

# 3. Calculate the Systematic Utility (V)
# This uses the model coefficients and the data matrix
df_math$V <- as.numeric(model.matrix(rum_mle) %*% coef(rum_mle))

# 4. Double check the columns are there
head(df_math[, c("trip_id", "park_name", "V")])



```

#### McFaddenR2

The mlogit gives NA so running manually to calculate McFadden to double check the R2

and it above .2 so that great

```{r }
# 1. Estimate your model (The "Full" model)
rum_full <- mlogit(choice ~ price | 1, data = df_rum)

# 2. Estimate the "Null" model (Intersects only, no price)
# Use '0' for the first part of the formula to remove choice-specific variables
rum_null <- mlogit(choice ~ 0 | 1, data = df_rum)

# 3. Use the lrtest function to get your Chi-square and p-value
library(lmtest)
lr_test_result <- lrtest(rum_full, rum_null)
print(lr_test_result)

# 4. Calculate McFadden's R-squared manually
ll_full <- as.numeric(logLik(rum_full))
ll_null <- as.numeric(logLik(rum_null))
mcfadden_r2 <- 1 - (ll_full / ll_null)

print(paste("Manual McFadden R2:", round(mcfadden_r2, 4)))



```

#### Per Trip Value

```{r }


# Get the unique list of parks from your data
# We exclude "Stay_Home" because we aren't "closing" the option to stay home
all_parks <- unique(df_math$park_name)
all_parks <- all_parks[all_parks != "Stay_Home"]

library(purrr)

# 1. Define the list of parks
all_parks <- unique(df_math$park_name)
all_parks <- all_parks[all_parks != "Stay_Home"]

# 2. Run the function for all parks and combine results
# This replaces your 'for' loop entirely
rum_results_2018 <- map_df(all_parks, function(p) {
  
  # Run your function (the one that returns mean, sd, ci)
  stats <- calc_rum_welfare_manual(p)
  
  # Add the park name so we know which row is which
  stats$parkname <- p
  
  return(stats)
})

# 3. Reorder columns to put parkname first
rum_results_2018 <- rum_results_2018 %>% 
  select(parkname, mean_cv_rum, std_dev_rum, ci_lo_2.5_rum, ci_hi_97.5_rum, n)

print(rum_results_2018)

rum_results_2018 <- rum_results_2018 %>% 
  select(parkname, mean_cv_rum, std_dev_rum, ci_lo_2.5_rum, ci_hi_97.5_rum, n) %>%
  mutate(
    # Use mean_cv_rum because that is what your function produced
    rum_total_annual_loss = mean_cv_rum * n
  )

library(dplyr)
library(stringr)
park_number <- final_data_2018 %>%
  # Select only the two columns you need
  select(parkname, parkname_numeric) %>%
  # Keep only the unique combinations
  distinct() %>%
  # Sort them by the numeric ID for easier reading
  arrange(parkname_numeric)

welfare_summary=read_csv("welfare_summary_2018.csv")

welfare_summary=welfare_summary %>%
  mutate(
    # Extract only the digits from the "policy" string and make it an integer
    parkname_numeric = as.integer(str_extract(policy, "\\d+"))
  ) %>%
  # 2. Merge with your park_number key (created in the previous step)
  left_join(park_number, by = "parkname_numeric") %>%
  # 3. Clean up: Move parkname to the front and remove the "policy" column
  select(parkname, parkname_numeric, mean, std.dev, everything(), -policy)



```

#### Comparison MDCEV

So lets look at the comparison of MDCEV which will give us a value for the closure and compare it to the per trip value gives us for a site closure.

```{r}
final_comparison_2018 <- rum_results_2018 %>%
  select(parkname, rum_mean_per_trip=mean_cv_rum, std_dev_rum, ci_lo_2.5_rum, ci_hi_97.5_rum, rum_total_annual_loss) %>%
  left_join(welfare_summary, by = "parkname") %>%
  rename(mdcev_mean_per_person = mean) %>%
  mutate(
    # 1. RUM Annual = (Mean per Trip) * (Total Number of Trips)
    rum_annual_total = rum_mean_per_trip * sum(final_data_2019$VisitCount),
    
    # 2. MDCEV Annual = (Mean per Person) * (Number of Unique People)
    # We use n_distinct(id) to get exactly 1,380
    mdcev_annual_total = mdcev_mean_per_person * n_distinct(final_data_2019$id),
    
    # 3. The Gap
    difference = rum_annual_total - mdcev_annual_total
  )
```

### Loop

I have each year thats basically copied from 2018 so if there is a change from 2018 then it needs to be changed for every single one.

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(mlogit)
library(stringr)
library(readr)

years <- 2018:2023

final_comparison_all_years <- map_df(years, function(yr) {
  
  message(paste(">>> Processing Year:", yr))
  
  # 0. Get current year data
  current_data <- get(paste0("final_data_", yr))
  
  # 1. Create unique choice occasions for every visit
  data_rum_raw <- current_data %>%
    filter(VisitCount > 0) %>%
    uncount(VisitCount) %>%
    group_by(id) %>%
    mutate(choice_id = paste0(id, "_", row_number())) %>%
    ungroup() %>%
    select(choice_id, id, chosen_park = parkname) %>%
    expand_grid(parkname = unique(current_data$parkname)) %>%
    mutate(choice = (chosen_park == parkname)) %>%
    left_join(current_data %>% select(id, parkname, price, travel_distance_km), 
              by = c("id", "parkname"))

  # 2. Add Stay-at-Home Option
  outside_home <- data_rum_raw %>%
    distinct(choice_id, id) %>%
    mutate(parkname = "Stay_Home", choice = FALSE, price = 0, travel_distance_km = 0)
  
  data_rum_combined <- bind_rows(data_rum_raw, outside_home)

  # 3. Model Estimation - Pushed to Global for the function to see
  df_idx_local <- dfidx(data_rum_combined, idx = list("choice_id", "parkname"), shape = "long")
  rum_mle <<- mlogit(choice ~ price | 1, data = df_idx_local)
  
  # 4. Prepare df_math - Pushed to Global for the function to see
  # This is the step that was likely missing or stale in your previous run
  df_math_local <- as.data.frame(df_idx_local)
  df_math_local$trip_id   <- as.character(idx(df_idx_local, 1))
  df_math_local$park_name <- as.character(idx(df_idx_local, 2))
  
  # Calculate V (Systematic Utility) using the CURRENT year's model
  df_math_local$V <- as.numeric(model.matrix(rum_mle) %*% coef(rum_mle))
  
  # Update Global df_math so calc_rum_welfare_manual uses the NEW year's V and trip_ids
  df_math <<- df_math_local

  # 5. RUM Welfare Calculations
  all_parks_list <- unique(df_math$park_name)
  all_parks_list <- all_parks_list[all_parks_list != "Stay_Home"]
  
  rum_results <- map_df(all_parks_list, function(p) {
    stats <- calc_rum_welfare_manual(p) 
    stats$parkname <- p
    return(stats)
  })

  # 6. Process Welfare Summary CSV
  welfare_summary_yr <- read_csv(paste0("welfare_summary_", yr, ".csv")) %>%
    mutate(parkname_numeric = as.integer(str_extract(policy, "\\d+"))) %>%
    left_join(
      current_data %>% select(parkname, parkname_numeric) %>% distinct(), 
      by = "parkname_numeric"
    ) %>%
    rename(mdcev_mean_per_person = mean)

  # 7. Final Join for the specific year
  year_comparison <- rum_results %>%
    mutate(rum_total_annual_loss = mean_cv_rum * n) %>%
    select(parkname, rum_mean_per_trip = mean_cv_rum, std_dev_rum, 
           ci_lo_2.5_rum, ci_hi_97.5_rum, rum_total_annual_loss) %>%
    left_join(welfare_summary_yr, by = "parkname") %>%
    mutate(
      year = yr,
      rum_annual_total = rum_mean_per_trip * sum(current_data$VisitCount, na.rm = TRUE),
      mdcev_annual_total = mdcev_mean_per_person * n_distinct(current_data$id),
      difference = rum_annual_total - mdcev_annual_total
    ) %>%
    select(year, parkname, rum_mean_per_trip, mdcev_mean_per_person, 
           rum_annual_total, mdcev_annual_total, difference, everything(), -policy)

  return(year_comparison)
})

write_csv(final_comparison_all_years,file="rum_combined_loop.csv")
```

# Results comparison

```{r}
combined_welfare=read_csv("rum_combined_totalWF.csv")
print(combined_welfare)

```

This would be the estimate for a closer of each park.

If MDCEV is the mean for **per person** and RUM mean is **per trip**, then to compare them by looking at the **Total Annual Welfare Loss** for each park:

$Total\ MDCEV\ Loss = (MDCEV\ MeanPerson) \times (Total\ Number\ of\ People)$

$Total\ RUM\ Loss = (RUM\ Mean Trip) \times (Total\ Number\ of\ Trips)$

```{r}

library(ggplot2)



# Prepare the data into a long format for the two models
plot_data_long <- combined_welfare %>%
  select(year, parkname, rum_annual_total, mdcev_annual_total) %>%
  pivot_longer(cols = c(rum_annual_total, mdcev_annual_total), 
               names_to = "Model", 
               values_to = "Annual_Loss") %>%
  mutate(Model = ifelse(Model == "rum_annual_total", "RUM", "MDCEV"),
         # Convert to absolute values (positive dollars) for easier reading
         Annual_Loss_Abs = abs(Annual_Loss))

ggplot(plot_data_long, aes(x = as.factor(year), y = Annual_Loss_Abs, fill = Model)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21, outlier.fill = "white") +
  scale_y_log10(labels = scales::dollar_format()) +
  labs(title = "Distribution of Annual Welfare Losses Across Parks",
       subtitle = "Comparing RUM vs MDCEV Park-Level Estimates (2018-2023)",
       x = "Year", 
       y = "Annual Welfare Loss per Park (Log Scale)") +
  theme_minimal() +
  scale_fill_manual(values = c("MDCEV" = "#2c3e50", "RUM" = "#e74c3c"))

```

## Comparison Individual

```{r}


# Create unique choice occasions for every visit
data_rum <- final_data_2018 %>%
  filter(VisitCount > 0) %>%
  uncount(VisitCount) %>%
  group_by(id) %>%
  mutate(choice_id = paste0(id, "_", row_number())) %>%
  ungroup() %>%
  select(choice_id, id, chosen_park = parkname) %>%
  expand_grid(parkname = unique(final_data_2018$parkname)) %>%
  mutate(choice = (chosen_park == parkname)) %>%
  left_join(final_data_2018 %>% select(id, parkname, price, travel_distance_km,income,zip), 
            by = c("id", "parkname"))

# Create Stay at Home

# Create the Stay-at-Home option for every choice occasion
outside_home <- data_rum %>%
  distinct(choice_id, id,income,zip) %>%
  mutate(
    parkname = "Stay_Home",
    choice = FALSE,      
    price = 0,
    travel_distance_km = 0
  )

# Combine with your existing data
data_rum<- bind_rows(data_rum, outside_home)

# Re-format for mlogit
df_rum <- dfidx(data_rum, 
                        idx = list("choice_id", "parkname"), 
                        shape = "long")

# Simple Multinomial Logit
rum_mle <- mlogit(choice ~ price  | 1, data = df_rum)
summary(rum_mle)




# 1. Generate the utilities for ALL alternatives (38,786 rows)
V_all <- as.numeric(model.matrix(rum_mle) %*% coef(rum_mle))

# 2. Extract the long-format model frame and attach V
df_long_v <- as.data.frame(model.frame(rum_mle))
df_long_v$V <- V_all
df_long_v$choice_id <- idx(rum_mle, 1)
df_long_v$parkname  <- idx(rum_mle, 2)

# 3. Get metadata (id and zip) to ensure we don't lose them
# We grab this from your 'data_rum' object
metadata <- data_rum %>% distinct(choice_id, id, zip)

# 4. Define the list of sites (excluding Stay_Home)
sites_to_test <- unique(df_long_v$parkname)
sites_to_test <- sites_to_test[sites_to_test != "Stay_Home"]

# 5. Extract Price Coefficient
beta_p <- abs(coef(rum_mle)["price"])

library(purrr)

# Use map_dfr to loop through sites and bind results into one dataframe
welfare_2018_rum_ind <- map_dfr(sites_to_test, function(s) {
    
    df_long_v %>%
        group_by(choice_id) %>%
        summarise(
            # Welfare change = [ln(sum exp V_all) - ln(sum exp V_minus_s)] / beta_p
            ls_all = log(sum(exp(V))),
            ls_minus_s = log(sum(exp(V[parkname != s]))),
            .groups = "drop"
        ) %>%
        mutate(
            site = s,
            trip_welfare = (ls_all - ls_minus_s) / beta_p
        ) %>%
        # Join metadata to get 'id' and 'zip'
        left_join(metadata, by = "choice_id") %>%
        # Sum up welfare across all trips for each person
        group_by(id, zip, site) %>%
        summarise(mean_value = sum(trip_welfare), .groups = "drop")
})

# Final view
head(welfare_2018_rum_ind)
```

From here lets compare the same hicksan Consumer Surplus the MDCEV is calculating the change from a loss of a site and holding utility constant

Where as the Rum utility we are measuring the marshallian where income is held constant and the indirect utility is the theory.

The measurement is higher when we looked at per site but is it the same if we look at individual? This is the next section.

```{r}
welfare_2018_ind <- read_csv("welfare_2018_ind.csv")


welfare_2018_rum_ind=merge(park_number,welfare_2018_rum_ind, by.x=c("parkname"), by.y =c("site") )

welfare_2018_ind <- read_csv("welfare_2018_ind.csv")
colnames(welfare_2018_rum_ind)[5]="mean_value_rum"
welfare_2018_ind_combined=merge(welfare_2018_rum_ind,welfare_2018_ind, by.x=c("parkname_numeric","id","zip"), by.y =c("site","id","zip") )
```

# Calculating Cumulative Value

We can think of the two as RUM gets us an idea of charging for a trip while the MDCEV gives us the total value to think about its

```{r}


library(ggplot2)
library(dplyr)
library(tidyr)

#"Staple" and "Satiating" parks for 2023
satiation_ranking <- combined_welfare %>%
  filter(year == 2023) %>%
  mutate(satiation_index = abs(mdcev_mean_per_person / rum_mean_per_trip)) %>%
  arrange(satiation_index)

#two extremes
staple_park <- satiation_ranking %>% slice_max(satiation_index, n = 1)
satiating_park <- satiation_ranking %>% slice_min(satiation_index, n = 1)

# Function to generate cumulative data
generate_cum_data <- function(park_row, label, max_trips = 30) {
  trip_seq <- 1:max_trips
  val_per_trip <- abs(park_row$rum_mean_per_trip)
  val_total_person <- abs(park_row$mdcev_mean_per_person)
  
  # RUM: Linear growth
  rum_cum <- trip_seq * val_per_trip
  
  # MDCEV: Logarithmic growth constrained to the 'Total Annual' result
  # We assume the MDCEV total is reached at roughly the 'average' frequency
  mdcev_cum <- val_total_person * (log10(trip_seq + 1) / log10(max_trips + 1))
  
  data.frame(
    Trip = rep(trip_seq, 2),
    Surplus = c(rum_cum, mdcev_cum),
    Model = rep(c("RUM (Trip-Based)", "MDCEV (Person-Based)"), each = max_trips),
    Park_Type = label,
    Park_Name = park_row$parkname
  )
}

# Combine data for both parks
plot_data_combined <- rbind(
  generate_cum_data(staple_park, "Staple Park (Low Satiation)"),
  generate_cum_data(satiating_park, "Satiating Park (High Satiation)")
)

# Create the Comparison Plot
ggplot(plot_data_combined, aes(x = Trip, y = Surplus, fill = Model, color = Model)) +
  geom_area(alpha = 0.3, position = "identity") +
  geom_line(size = 1) +
  facet_wrap(~Park_Type + Park_Name, scales = "free_y") +
  scale_y_continuous(labels = scales::dollar) +
  scale_fill_manual(values = c("MDCEV (Person-Based)" = "#2c3e50", "RUM (Trip-Based)" = "#e74c3c")) +
  scale_color_manual(values = c("MDCEV (Person-Based)" = "#2c3e50", "RUM (Trip-Based)" = "#e74c3c")) +
  labs(
    title = "Satiation Comparison: How Model Choice Diverges with Frequency",
    subtitle = "The 'Red Gap' represents surplus RUM creates that MDCEV assumes is lost to satiation.",
    x = "Number of Trips per Year",
    y = "Cumulative Consumer Surplus ($)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", strip.text = element_text(face="bold"))
```

## 52-weeks

Testing the idea of putting more choosing stay at home... Increases WTP. Problem here is final data doesn't have day so maybe this is not a good assumption.

```{r}
library(dplyr)
library(tidyr)

# 1. Get all unique users and all unique parks from current data
unique_users <- final_data_2018 %>% distinct(id, zip, income)
all_parks <- unique(final_data_2018$parkname)

# 2. Create the "52 Week" Skeleton
# This creates 52 rows for every single person in dataset
df_52_weeks <- unique_users %>%
  expand_grid(week = 1:52) %>%
  mutate(choice_id = paste0(id, "_w", week))

# 3. Identify the ACTUAL trips taken (from reservation data)
# need to assign these to specific weeks. 
# if no exact dates, we distribute them randomly or to the first few weeks.
actual_trips <- data_rum %>%
  filter(choice == TRUE & parkname != "Stay_Home") %>%
  group_by(id) %>%
  mutate(week = row_number()) %>% # Assigns trip 1 to week 1, trip 2 to week 2
  ungroup() %>%
  select(id, week, chosen_park = parkname)

# 4. Join the actual trips into the 52-week skeleton
df_expanded_rum <- df_52_weeks %>%
  left_join(actual_trips, by = c("id", "week")) %>%
  # If 'chosen_park' is NA, it means they stayed home that week
  mutate(chosen_park = ifelse(is.na(chosen_park), "Stay_Home", chosen_park))
# 5. Expand to full choice sets (Parks + Stay_Home) for every week
df_rum_final <- df_expanded_rum %>%
  expand_grid(parkname = c(all_parks, "Stay_Home")) %>%
  mutate(choice = (chosen_park == parkname))

# 6. Re-attach Prices and Distance
# Note: Price for 'Stay_Home' is always 0.
df_rum_final <- df_rum_final %>%
  left_join(final_data_2018 %>% select(id, parkname, price, travel_distance_km), 
            by = c("id", "parkname")) %>%
  mutate(
    price = ifelse(parkname == "Stay_Home", 0, price),
    travel_distance_km = ifelse(parkname == "Stay_Home", 0, travel_distance_km)
  )

# 7. Re-format for mlogit
library(mlogit)
df_idx_52 <- dfidx(df_rum_final, 
                   idx = list("choice_id", "parkname"), 
                   shape = "long")

rum_mle_52 <- mlogit(choice ~ price | 1 |income , data = df_idx_52)

summary(rum_mle_52)
```

#### 

```{r}
# Calculate the 'Logsum' (Total Utility) for the status quo
logsum_base <- logsum(rum_mle_52)

# 1. Convert df_rum to a standard data frame for our math
df_math <- as.data.frame(df_idx_52)

# 2. Extract the IDs and Park names from the 'idx' column
# idx(df_rum, 1) gets the 'choice_id'
# idx(df_rum, 2) gets the 'parkname'
df_math$trip_id   <- as.character(idx(df_idx_52, 1))
df_math$park_name <- as.character(idx(df_idx_52, 2))

# 3. Calculate the Systematic Utility (V)
# This uses the model coefficients and the data matrix
df_math$V <- as.numeric(model.matrix(rum_mle_52) %*% coef(rum_mle_52))

# 4. Double check the columns are there
head(df_math[, c("trip_id", "park_name", "V")])


# Get the unique list of parks from your data
# We exclude "Stay_Home" because we aren't "closing" the option to stay home
all_parks <- unique(df_math$park_name)
all_parks <- all_parks[all_parks != "Stay_Home"]

library(purrr)

# 1. Define the list of parks
all_parks <- unique(df_math$park_name)
all_parks <- all_parks[all_parks != "Stay_Home"]

# 2. Run the function for all parks and combine results
# This replaces your 'for' loop entirely
rum_results_2018_52 <- map_df(all_parks, function(p) {
  
  # Run your function (the one that returns mean, sd, ci)
  stats <- calc_rum_welfare_manual(p)
  
  # Add the park name so we know which row is which
  stats$parkname <- p
  
  return(stats)
})

# 3. Reorder columns to put parkname first
rum_results_2018_52 <- rum_results_2018_52 %>% 
  select(parkname, mean_cv_rum, std_dev_rum, ci_lo_2.5_rum, ci_hi_97.5_rum, n)

print(rum_results_2018_52)

rum_results_2018_52 <- rum_results_2018_52 %>% 
  select(parkname, mean_cv_rum, std_dev_rum, ci_lo_2.5_rum, ci_hi_97.5_rum, n) %>%
  mutate(
    # Use mean_cv_rum because that is what your function produced
    rum_total_annual_loss = mean_cv_rum * n
  )

library(dplyr)
library(stringr)
park_number <- final_data_2018 %>%
  # Select only the two columns you need
  select(parkname, parkname_numeric) %>%
  # Keep only the unique combinations
  distinct() %>%
  # Sort them by the numeric ID for easier reading
  arrange(parkname_numeric)

welfare_summary=read_csv("welfare_summary_2018.csv")

welfare_summary=welfare_summary %>%
  mutate(
    # Extract only the digits from the "policy" string and make it an integer
    parkname_numeric = as.integer(str_extract(policy, "\\d+"))
  ) %>%
  # 2. Merge with your park_number key (created in the previous step)
  left_join(park_number, by = "parkname_numeric") %>%
  # 3. Clean up: Move parkname to the front and remove the "policy" column
  select(parkname, parkname_numeric, mean, std.dev, everything(), -policy)

final_comparison_2018_52 <- rum_results_2018_52 %>%
  select(parkname, rum_mean_per_trip=mean_cv_rum, std_dev_rum, ci_lo_2.5_rum, ci_hi_97.5_rum, rum_total_annual_loss) %>%
  left_join(welfare_summary, by = "parkname") %>%
  rename(mdcev_mean_per_person = mean) %>%
  mutate(
    # 1. RUM Annual = (Mean per Trip) * (Total Number of Trips)
    rum_annual_total = rum_mean_per_trip * sum(final_data_2019$VisitCount),
    
    # 2. MDCEV Annual = (Mean per Person) * (Number of Unique People)
    mdcev_annual_total = mdcev_mean_per_person * n_distinct(final_data_2018$id),
    
    # 3. The Gap
    difference = rum_annual_total - mdcev_annual_total
  )

```

### Loop code

This just simiplifes the

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(mlogit)
library(stringr)
library(readr)

years <- 2018:2023

final_comparison_52 <- map_df(years, function(yr) {
    
    message(paste(">>> Processing Year (52-week model):", yr))
    
    # 0. Get current year data
    current_data <- get(paste0("final_data_", yr))
    
    # 1. Skeleton Setup
    unique_users <- current_data %>% distinct(id, zip, income)
    all_parks <- unique(current_data$parkname)
    
    df_52_weeks <- unique_users %>%
        expand_grid(week = 1:52) %>%
        mutate(choice_id = paste0(id, "_w", week))
    
    # Ensure actual_trips is specific to the current year's data
    actual_trips <- current_data %>% 
        filter(VisitCount > 0) %>% # Using current_data to find where trips actually happened
        uncount(VisitCount) %>%
        group_by(id) %>%
        mutate(week = row_number()) %>%
        filter(week <= 52) %>% # Safety check for skeleton bounds
        ungroup() %>%
        select(id, week, chosen_park = parkname)
    
    # 2. Build Choice Set
    df_rum_final <- df_52_weeks %>%
        left_join(actual_trips, by = c("id", "week")) %>%
        mutate(chosen_park = replace_na(chosen_park, "Stay_Home")) %>%
        expand_grid(parkname = c(all_parks, "Stay_Home")) %>%
        mutate(choice = (chosen_park == parkname)) %>%
        left_join(current_data %>% select(id, parkname, price, travel_distance_km), 
                  by = c("id", "parkname")) %>%
        mutate(across(c(price, travel_distance_km), ~replace_na(.x, 0)))
    
    # 3. Model Estimation - Update GLOBAL rum_mle_52
    df_idx_local <- dfidx(df_rum_final, idx = list("choice_id", "parkname"), shape = "long")
    rum_mle_52 <<- mlogit(choice ~ price | 1 , data = df_idx_local)
    
    # 4. Prepare df_math - Update GLOBAL df_math
    # This is the crucial step for your function to calculate correct logsums
    df_math_local <- as.data.frame(df_idx_local)
    df_math_local$trip_id   <- as.character(idx(df_idx_local, 1))
    df_math_local$park_name <- as.character(idx(df_idx_local, 2))
    
    # Recalculate V for the current year's model coefficients
    df_math_local$V <- as.numeric(model.matrix(rum_mle_52) %*% coef(rum_mle_52))
    
    # Push to Global
    df_math <<- df_math_local
    
    # 5. RUM Welfare Calculations
    parks_to_test <- all_parks[all_parks != "Stay_Home"]
    
    rum_results <- map_df(parks_to_test, function(p) {
        # This now finds the UPDATED rum_mle_52 and df_math
        stats <- calc_rum_welfare_manual(p) 
        stats$parkname <- p
        return(stats)
    })
    
    # 6. Process Welfare Summary CSV
    welfare_summary_yr <- read_csv(paste0("welfare_summary_", yr, ".csv")) %>%
        mutate(parkname_numeric = as.integer(str_extract(policy, "\\d+"))) %>%
        left_join(current_data %>% select(parkname, parkname_numeric) %>% distinct(), 
                  by = "parkname_numeric") %>%
        rename(mdcev_mean_per_person = mean)
    
    # 7. Final Join and Metrics
    year_comparison <- rum_results %>%
        mutate(rum_total_annual_loss = mean_cv_rum * n) %>%
        select(parkname, rum_mean_per_trip = mean_cv_rum, rum_total_annual_loss,
               std_dev_rum, ci_lo_2.5_rum, ci_hi_97.5_rum) %>%
        left_join(welfare_summary_yr, by = "parkname") %>%
        mutate(
            year = yr,
            mdcev_annual_total = mdcev_mean_per_person * n_distinct(current_data$id),
            rum_annual_total = rum_mean_per_trip * sum(current_data$VisitCount, na.rm = TRUE),
            difference = rum_annual_total - mdcev_annual_total
        ) %>%
        select(year, parkname, rum_mean_per_trip, mdcev_mean_per_person, 
               rum_annual_total, mdcev_annual_total, difference, everything(), -policy)
    
    return(year_comparison)
})

# View the final combined dataset
print(final_comparison_52)


```

# Compare Total

The MDVEC gives total surplus from choke price for each site

Here this shows how dropping out a site in the RUM and recalculating the CS using the logsum process compares to this same process. One we look at just a simple rum and then we expand the RUM to include a choice every week 52 choice sets for each id. For the ones we don't observe during the week we assume they choose stay at home option.

This is to try to get a comparision of the Hiskan composite good.

Interesting part really is the over and under estimation where as the MDCEV provides a middle ground.. What do you think Anders?

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

# 1. Merge and Reshape
plot_data <- final_comparison_all_years %>%
  select(year, parkname, RUM_Standard = rum_annual_total, MDCEV = mdcev_annual_total) %>%
  inner_join(
    select(final_comparison_52, year, parkname, RUM_52_Week = rum_annual_total),
    by = c("year", "parkname")
  ) %>%
  pivot_longer(cols = c(RUM_Standard, RUM_52_Week, MDCEV), 
               names_to = "Model", values_to = "Loss") %>%
  mutate(Loss_Abs = abs(Loss))

# 2. Plot with Times New Roman
ggplot(plot_data, aes(x = factor(year), y = Loss_Abs, fill = Model)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1) +
  scale_y_log10(labels = scales::dollar_format()) +
  scale_fill_manual(values = c("#2c3e50", "#e74c3c", "#3498db")) +
  labs(title = "Predicted Annual Losses across 22 Site Closures",
       subtitle = "Comparing Range from Standard RUM, 52-Week RUM, and MDCEV (2018-2023)",
       x = "Year", y = "Annual Welfare Loss") +
  # 'serif' maps to Times New Roman
  theme_minimal(base_family = "serif") + 
  theme(legend.position = "bottom",
        text = element_text(size = 12),
        plot.title = element_text(face = "bold"))


```